---
title: "3 - What makes a model? - Classwork"
subtitle: "Machine learning with tidymodels"
editor_options: 
  chunk_output_type: console
---

We recommend restarting R between each slide deck!

## Setup

Setup from deck 2

```{r}
library(tidymodels)
library(modeldatatoo)

taxi <- data_taxi()

taxi <- taxi %>%
  mutate(month = factor(month, levels = c("Jan", "Feb", "Mar", "Apr"))) %>% 
  select(-c(id, duration, fare, tolls, extras, total_cost, payment_type)) %>% 
  drop_na()

set.seed(123)

taxi_split <- initial_split(taxi, prop = 0.8, strata = tip)
taxi_train <- training(taxi_split)
taxi_test <- testing(taxi_split)
```

## Your turn

How do you fit a linear model in R?

How many different ways can you think of?

Discuss with your neighbor!

## To specify a model

```{r}
# Model
linear_reg()

# Engine
linear_reg() %>%
  set_engine("glmnet")

# Mode - Some models have a default mode, others don't
decision_tree() %>% 
  set_mode("regression")
```

## Your turn

Edit the chunk below to use a different model!

All available models are listed at <https://www.tidymodels.org/find/parsnip/> 

```{r tree_spec}
tree_spec <- decision_tree() %>% 
  set_mode("classification")

tree_spec
```

## A model workflow

```{r}
tree_spec <-
  decision_tree() %>% 
  set_mode("classification")
```

Fit with parsnip:

```{r}
tree_spec %>% 
  fit(tip ~ ., data = taxi_train) 
```

Fit with a workflow:

```{r}
workflow() %>%
  add_formula(tip ~ .) %>%
  add_model(tree_spec) %>%
  fit(data = taxi_train) 
```

"Shortcut" by specifying the preprocessor and model spec directly in the `workflow()` call:

```{r}
workflow(tip ~ ., tree_spec) %>% 
  fit(data = taxi_train) 
```

## Your turn

Edit the chunk below to make a workflow with your own model of choice!

```{r tree_wflow}
tree_spec <-
  decision_tree() %>% 
  set_mode("classification")

tree_wflow <- workflow() %>%
  add_formula(tip ~ .) %>%
  add_model(tree_spec)

tree_wflow
```

## Predict with your model

```{r}
tree_fit <-
  workflow(tip ~ ., tree_spec) %>% 
  fit(data = taxi_train) 
```

## Your turn

What do you get from running the following code? What do you notice about the structure of the result?

```{r}
predict(tree_fit, new_data = taxi_test)
```

## Your turn

What do you get from running the following code? How is `augment()` different from `predict()`?

```{r}
augment(tree_fit, new_data = taxi_test)
```

## Understand your model

```{r}
library(rpart.plot)

tree_fit %>%
  extract_fit_engine() %>%
  rpart.plot(roundint = FALSE)
```

## Your turn

Try extracting the model engine object from your fitted workflow!

```{r}
### Your code here

```

What kind of object is it? What can you do with it?

⚠️ Never `predict()` with any extracted components!

You can also read the documentation for object extraction:
https://workflows.tidymodels.org/reference/extract-workflow.html

## Your turn

Explore how you might deploy your `tree_fit` model using vetiver.

```{r vetiver, eval = FALSE}
library(vetiver)
library(plumber)

# Create a vetiver model object
v <- vetiver_model(tree_fit, "taxi_tips")
v
```

```{r, eval = FALSE}
# Create a predictable Plumber API
pr <- pr() %>%
  vetiver_api(v)

pr
```

```{r, eval = FALSE}
# Run the API server in a new window
pr_run(pr)
```
