---
title: "3 - Grid Search via Racing"
subtitle: "Advanced tidymodels"
format:
  revealjs: 
    slide-number: true
    footer: <https://workshops.tidymodels.org>
    include-before-body: header.html
    include-after-body: footer-annotations.html
    theme: [default, tidymodels.scss]
    width: 1280
    height: 720
knitr:
  opts_chunk: 
    echo: true
    collapse: true
    comment: "#>"
    fig.path: "figures/"
---

```{r setup}
#| include: false
#| file: setup.R
```

```{r more-setup}
#| include: false
library(modeldatatoo)
library(probably)
library(countdown)
library(finetune)

cores <- parallelly::availableCores(logical = FALSE)
cl <- parallel::makePSOCKcluster(cores)
doParallel::registerDoParallel(cl)

options(width = 200)

ggplot2::theme_set(ggplot2::theme_bw())
```

## Previously - Setup `r hexes("tidymodels")`

:::: {.columns}

::: {.column width="40%"}

```{r}
#| label: tune-startup
library(tidymodels)
library(modeldatatoo)
library(textrecipes)
library(bonsai)

# Max's usual settings: 
tidymodels_prefer()
theme_set(theme_bw())
options(
  pillar.advice = FALSE, 
  pillar.min_title_chars = Inf
)

reg_metrics <- metric_set(mae, rsq)
```

:::

::: {.column width="60%"}

```{r}
#| label: data-import
set.seed(295)
hotel_rates <- 
  data_hotel_rates() %>% 
  sample_n(5000) %>% 
  arrange(arrival_date) %>% 
  select(-arrival_date_num, -arrival_date) %>% 
  mutate(
    company = factor(as.character(company)),
    country = factor(as.character(country)),
    agent = factor(as.character(agent))
  )
```


:::

::::


## Previously - Data Usage `r hexes("rsample")`

```{r}
#| label: hotel-split
set.seed(4028)
hotel_split <-
  initial_split(hotel_rates, strata = avg_price_per_room)

hotel_tr <- training(hotel_split)
hotel_te <- testing(hotel_split)

set.seed(472)
hotel_rs <- vfold_cv(hotel_tr, strata = avg_price_per_room)

```

## Previously - Boosting Model `r hexes("recipes", "textrecipes", "tune", "bonsai")`

```{r}
#| label: setup-lgbm
hotel_rec <-
  recipe(avg_price_per_room ~ ., data = hotel_tr) %>%
  step_YeoJohnson(lead_time) %>%
  step_dummy_hash(agent,   num_terms = tune("agent hash")) %>%
  step_dummy_hash(company, num_terms = tune("company hash")) %>%
  step_zv(all_predictors())

lgbm_spec <- 
  boost_tree(trees = tune(), learn_rate = tune(), min_n = tune()) %>% 
  set_mode("regression") %>% 
  set_engine("lightgbm")

lgbm_wflow <- workflow(hotel_rec, lgbm_spec)

lgbm_param <-
  lgbm_wflow %>%
  extract_parameter_set_dials() %>%
  update(`agent hash`   = num_hash(c(3, 8)),
         `company hash` = num_hash(c(3, 8)))
```


## Making Grid Search More Efficient

In the last section, we evaluated 250 models (25 candidates times 10 resamples).

We can make this go faster using parallel processing. 

Also, for some models, we can _fit_ far fewer models than the number that are being evaluated. 
 
 * For boosting, a model with `X` trees can often predict on candidates with less than `X` trees. 
 
Both of these methods can lead to enormous speed-ups. 


## Model Racing 

[_Racing_](https://scholar.google.com/scholar?hl=en&as_sdt=0%2C7&q=+Hoeffding+racing) is an old tool that we can use to go even faster. 

1. Evaluate all of the candidate models but only for a few resamples. 
1. Determine which candidates have a low probability of being selected.
1. Eliminate poor candidates.
1. Repeat with next resample (until no more resamples remain) 

This can result in fitting a small number of models. 


## Discarding Candidates

How do we eliminate tuning parameter combinations? 

There are a few methods to do so. We'll use one based on analysis of variance (ANOVA). 

_However_... there is typically a large difference between resamples in the results. 

## Resampling Results (Non-Racing)

:::: {.columns}

::: {.column width="50%"}


```{r}
#| label: race-data-comp
#| echo: false
#| out-width: 60%
#| fig-align: center
#| fig-width: 5
#| fig-height: 5

# Simulate some resamples for candidate models
num_resamples <- 10
race_example <- crossing(model = paste("candidate", 1:2), resample = 1:num_resamples)
candidates <- tibble(model = paste("candidate", 1:2), mean = c(4, 2))
set.seed(937)
resamples <- tibble(resample = 1:num_resamples, effect = rnorm(num_resamples, sd = 2))
race_example <-
  race_example %>%
  full_join(candidates, by = "model") %>%
  full_join(resamples, by = "resample") %>%
  mutate(
    error = mean + effect + rnorm(n(), sd = 1),
    resample = format(resample)
  ) %>%
  select(-mean, -effect)


race_table <- 
  race_example %>%
  mutate(model = gsub(" ", "_", model)) %>%
  pivot_wider(id_cols = c(resample), names_from = model, values_from = error)

race_cor <- cor(race_table[, -1], method = "spearman")[1,2]

get_ci <- function(ind) {
  dat <- race_table[1:ind,]
  t_test <- t.test(dat$candidate_1, dat$candidate_2, paired = TRUE)
  t_test <- tidy(t_test, conf)
  t_test %>% 
    select(difference = estimate, lower = conf.low, upper = conf.high) %>% 
    mutate(`number of resamples` = ind)
}

race_ci <- map_dfr(2:10, get_ci) 

```

Here are some realistic (but simulated) examples of two candidate models. 

An error estimate is measured for each of 10 resamples. 

 - The lines connect resamples. 

There is usually a significant resample-to-resample effect (rank corr: `r round(race_cor, 2)`). 

:::

::: {.column width="50%"}


```{r}
#| label: race-data
#| echo: false
#| out-width: 100%
#| fig-align: center
#| fig-width: 5
#| fig-height: 5


race_example %>%
  ggplot(aes(x = model, y = error, group = resample, col = factor(resample))) +
  geom_point(show.legend = FALSE) +
  geom_line(alpha = 1 / 2, show.legend = FALSE) +
  labs(x = NULL, y = "model error")
```

:::

::::


## Are Candidates Different?

One way to evaluate these models is to do a paired t-test
 
 - or a t-test on their differences matched by resamples

With $n = 10$ resamples, the confidence interval is (`r signif(race_ci$lower[9], 2)`, `r signif(race_ci$upper[9], 2)`), indicating that candidate number 2 has smaller error. 

What if we were to compare each model candidate to the current best at each resample? 

One shows superiority when `r which.max(race_ci$lower > 0)` resamples have been evaluated.


## Evaluating Differences in Candidates

```{r}
#| label: race-ci
#| echo: false
#| out-width: 70%
#| fig-align: center
#| fig-width: 6
#| fig-height: 4.25


race_ci %>% 
  ggplot(aes(`number of resamples`, difference)) + 
  geom_point() +
  geom_hline(yintercept = 0, col = "red", lty = 2) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 1 / 10, alpha = 0.5) +
  labs(y = "error difference from best") + 
  scale_x_continuous(breaks = pretty_breaks())
```

## Interim Analysis of Results

One version of racing uses a _mixed model ANOVA_ to construct one-sided confidence intervals for each candidate versus the current best. 

Any candidates whose bound does not include zero are discarded.  [Here](https://www.tmwr.org/race_results.mp4) is an animation.

The resamples are analyzed in a random order.

<br>

[Kuhn (2014)](https://arxiv.org/abs/1405.6974) has examples and simulations to show that the method works. 

The [finetune](https://finetune.tidymodels.org/) package has functions `tune_race_anova()` and `tune_race_win_loss()`. 


## Racing `r hexes("recipes", "textrecipes", "tune", "bonsai", "finetune")`


```{r } 
#| label: lgb-grid-race
#| cache: false
#| code-line-numbers: "1-8|"
# Let's use a larger grid
set.seed(8945)
lgbm_grid <- 
  lgbm_param %>% 
  grid_latin_hypercube(size = 50)

library(finetune)

set.seed(9)
lgbm_race_res <-
  lgbm_wflow %>%
  tune_race_anova(
    resamples = hotel_rs,
    grid = lgbm_grid, 
    metrics = reg_metrics
  )
```

The syntax and helper functions are extremely similar to those shown for `tune_grid()`. 


## Racing Results `r hexes("tune")`

```{r}
#| label: best-race

show_best(lgbm_race_res, metric = "mae")
```

## Racing Results `r hexes("finetune")`

:::: {.columns}

::: {.column width="50%"}
Only `r sum(map_int(lgbm_race_res$.metrics, ~ nrow(.x) / 2))` models were fit (out of `r nrow(lgbm_race_res) * nrow(lgbm_grid)`). 

`select_best()` never considers candidate models that did not get to the end of the race. 

There is a helper function to see how candidate models were removed from consideration. 

:::

::: {.column width="50%"}


```{r}
#| label: plot-race
#| out-width: 100%
#| fig-align: center
#| fig-width: 5
#| fig-height: 3.8

plot_race(lgbm_race_res) + 
  scale_x_continuous(breaks = pretty_breaks())
```

:::

::::


## Your turn {transition="slide-in"}

- *Run `tune_race_anova()` with a different seed.*
- *Did you get the same or similar results?*


```{r}
#| label: racing-repeat
#| echo: false
countdown(minutes = 10, id = "racing-repeat")
```

```{r}
#| label: teardown
#| echo: false

parallel::stopCluster(cl)
```
