---
title: "1 - Introduction"
subtitle: "Advanced tidymodels"
format:
  revealjs: 
    slide-number: true
    footer: <https://workshops.tidymodels.org>
    include-before-body: header.html
    include-after-body: footer-annotations.html
    theme: [default, tidymodels.scss]
    width: 1280
    height: 720
knitr:
  opts_chunk: 
    echo: true
    collapse: true
    comment: "#>"
---

```{r setup}
#| include: false
#| file: setup.R
```

::: r-fit-text
Welcome!
:::

::: columns
::: {.column width="50%"}

<center>

### <i class="fa fa-wifi"></i>

Wi-Fi network name

`Posit Conf 2023`

</center>

:::

::: {.column width="50%"}

<center>

### <i class="fa fa-key"></i>

Wi-Fi password

`conf2023`

</center>

:::
:::

## Workshop policies

-   Please do not photograph people wearing red lanyards

-   There are gender-neutral bathrooms located are among the Grand Suite Bathrooms

-   There are two meditation/prayer rooms: Grand Suite 2A and 2B

-   A lactation room is located in Grand Suite 1

-   The meditation/prayer and lactation rooms are open  
    Sun - Tue 7:30am - 7:00pm, Wed 8:00am - 6:00pm

## Workshop policies

-   Please review the code of conduct and COVID policies, which apply to all workshops: <https://posit.co/code-of-conduct/.>

-   CoC site has info on how to report a problem (in person, email, phone)

## Who are you?

-   You can use the magrittr `%>%` or base R `|>` pipe

-   You are familiar with functions from dplyr, tidyr, ggplot2

-   You have exposure to basic statistical concepts

-   You do **not** need intermediate or expert familiarity with modeling or ML

-   You have used some tidymodels packages
 
-   You have some experience with evaluating statistical models using resampling techniques 

## Who are tidymodels?

-   Simon Couch
-   Hannah Frick
-   Emil Hvitfeldt
-   Max Kuhn

. . .

Ijeamaka Anyene (Day 1) and Edgar Ruiz (Day 2) are TAing!

. . .

Many thanks to Davis Vaughan, Julia Silge, David Robinson, Julie Jung, Alison Hill, and Desirée De Leon for their role in creating these materials!

## Asking for help

. . .

🟪 "I'm stuck and need help!"

. . .

🟩 "I finished the exercise"


## `r emo::ji("eyes")` {.annotation}

![](images/pointing.svg){.absolute top="0" right="0"}

## Tentative plan for this workshop

- Feature engineering with recipes
- Model optimization by tuning
  - Grid search
  - Racing
  - Iterative methods
- Extras (time permitting)
  - Effect encodings
  - A case study

##  {.center}

### Introduce yourself to your neighbors 👋

<br></br>

### <i class="fa fa-cloud"></i> Log in to Posit Cloud (free):

Check the workshop channel on Discord for the link!

## Let's install some packages

If you are using your own laptop instead of RStudio Cloud:

```{r load-pkgs}
#| eval: false

# Install the packages for the workshop
pkgs <- 
  c("bonsai", "doParallel", "embed", "finetune", "lightgbm", "lme4",
    "plumber", "probably", "ranger", "rpart", "rpart.plot", "rules",
    "splines2", "stacks", "text2vec", "textrecipes", "tidymodels", 
    "vetiver", "remotes")

install.packages(pkgs)
```

. . .

<br></br>

### <i class="fa fa-cloud"></i> Or log in to Posit Cloud

Link in our Discord channel!

## Hotel Data `r hexes("tidymodels", "dplyr")`

We'll use [data on hotels](https://www.sciencedirect.com/science/article/pii/S2352340918315191) to predict the cost of a room. 

The [data](https://modeldata.tidymodels.org/reference/hotel_rates.html) are in the modeldata package. We'll sample down the data and refactor some columns: 

:::: {.columns}

::: {.column width="40%"}

```{r}
#| label: tune-startup
library(tidymodels)

# Max's usual settings: 
tidymodels_prefer()
theme_set(theme_bw())
options(
  pillar.advice = FALSE, 
  pillar.min_title_chars = Inf
)
```

:::

::: {.column width="60%"}

```{r}
#| label: data-import
data(hotel_rates)
set.seed(295)
hotel_rates <- 
  hotel_rates %>% 
  sample_n(5000) %>% 
  arrange(arrival_date) %>% 
  select(-arrival_date) %>% 
  mutate(
    company = factor(as.character(company)),
    country = factor(as.character(country)),
    agent = factor(as.character(agent))
  )
```


:::

::::

## Hotel date columns

```{r}
#| labels: hotel-names

names(hotel_rates)
```


## Data splitting strategy

```{r}
#| label: spending-diagram
#| fig-align: "center"
#| out-width: "20%"
#| echo: false

knitr::include_graphics("images/initial-split.svg")
```


## Data Spending `r hexes("rsample")`

Let's split the data into a training set (75%) and testing set (25%) using stratification:

```{r}
#| label: hotel-split
set.seed(4028)
hotel_split <- initial_split(hotel_rates, strata = avg_price_per_room)

hotel_train <- training(hotel_split)
hotel_test <- testing(hotel_split)
```



## Your turn {transition="slide-in"}

Let's take some time and investigate the _training data_. The outcome is `avg_price_per_room`. 

Are there any interesting characteristics of the data?

```{r}
#| label: hotel-investigation-exercise
#| echo: false
countdown::countdown(minutes = 10, id = "hotel-investigation")
```

## Our versions

```{r pkg-list, echo = FALSE}
deps <- 
  c("bonsai", "doParallel", "embed", "finetune", "lightgbm", "lme4",
    "plumber", "probably", "ranger", "rpart", "rpart.plot", "rules",
    "splines2", "stacks", "text2vec", "textrecipes", "tidymodels", 
    "vetiver", "remotes")

loaded <- purrr::map(deps, ~ library(.x, character.only = TRUE))
excl <- c("iterators", "emo", "countdown", "stats", "graphics", 
          "grDevices", "utils", "datasets", "methods", "base", "forcats", 
          "infer", "foreach", "Matrix", "R6", "parallel", "devtools", "usethis")
loaded <- loaded[[length(loaded)]]
loaded <- loaded[!(loaded %in% excl)]
pkgs <- 
  sessioninfo::package_info(loaded, dependencies = FALSE) %>% 
  select(-date)
df <- tibble::tibble(
  package = pkgs$package,
  version = pkgs$ondiskversion
)

ids <- split(
  seq_len(nrow(df)), 
  ceiling(seq_len(nrow(df)) / ceiling(nrow(df) / 4))
)

column1 <- df %>%
  dplyr::slice(ids[[1]])

column2 <- df %>%
  dplyr::slice(ids[[2]])

column3 <- df %>%
  dplyr::slice(ids[[3]])

column4 <- df %>%
  dplyr::slice(ids[[4]])

quarto_info <- paste0("Quarto (", system("quarto --version", intern = TRUE), ")")
```

`r R.version.string`, `r quarto_info`

::: {.columns style="font-size:0.7em;"}
::: {.column width="25%"}
```{r}
#| echo: false
knitr::kable(column1)
```
:::

::: {.column width="25%"}
```{r}
#| echo: false
knitr::kable(column2)
```
:::

::: {.column width="25%"}
```{r}
#| echo: false
knitr::kable(column3)
```
:::

::: {.column width="25%"}
```{r}
#| echo: false
knitr::kable(column4)
```
:::
:::

