---
title: "Extras - Effect Encodings"
subtitle: "Advanced tidymodels"
format:
  revealjs: 
    slide-number: true
    footer: <https://workshops.tidymodels.org>
    include-before-body: header.html
    include-after-body: footer-annotations.html
    theme: [default, tidymodels.scss]
    width: 1280
    height: 720
knitr:
  opts_chunk: 
    echo: true
    collapse: true
    comment: "#>"
    fig.path: "figures/"
---

```{r}
#| label: setup
#| include: false
#| file: setup.R
```

```{r more-setup}
#| include: false
library(modeldatatoo)
library(probably)
library(countdown)
library(finetune)

cores <- parallelly::availableCores(logical = FALSE)
cl <- parallel::makePSOCKcluster(cores)
doParallel::registerDoParallel(cl)

options(width = 200)

ggplot2::theme_set(ggplot2::theme_bw())
```

## Previously - Setup

:::: {.columns}

::: {.column width="40%"}

```{r}
#| label: tune-startup
library(tidymodels)
library(modeldatatoo)
library(textrecipes)
library(bonsai)

# Max's usual settings: 
tidymodels_prefer()
theme_set(theme_bw())
options(
  pillar.advice = FALSE, 
  pillar.min_title_chars = Inf
)
```

:::

::: {.column width="60%"}

```{r}
#| label: data-import
set.seed(295)
hotel_rates <- 
  data_hotel_rates() %>% 
  sample_n(5000) %>% 
  arrange(arrival_date) %>% 
  select(-arrival_date_num, -arrival_date) %>% 
  mutate(
    company = factor(as.character(company)),
    country = factor(as.character(country)),
    agent = factor(as.character(agent))
  )
```


:::

::::


## Previously - Data Usage

```{r}
#| label: hotel-split
set.seed(4028)
hotel_split <-
  initial_split(hotel_rates, strata = avg_price_per_room)

hotel_tr <- training(hotel_split)
hotel_te <- testing(hotel_split)

set.seed(472)
hotel_rs <- vfold_cv(hotel_tr, strata = avg_price_per_room)
```


## What do we do with the agent and company data? 

There are `r length(unique(hotel_tr$agent))` unique agent values and `r length(unique(hotel_tr$company))` companies in our training set. How can we include this information in our model?

. . .

We could:

-   make the full set of indicator variables 😳

-   lump agents and companies that rarely occur into an "other" group

-   use [feature hashing](https://www.tmwr.org/categorical.html#feature-hashing) to create a smaller set of indicator variables

-   use effect encoding to replace the `agent` and `company` columns with the estimated effect of that predictor



```{r}
#| label: effects-calcs 
#| include: false

agent_stats <- 
  hotel_tr %>%
  group_by(agent) %>%
  summarize(
    ADR = mean(avg_price_per_room), 
    num_reservations = n(),
    .groups = "drop"
    ) %>%
  mutate(agent = reorder(agent, ADR))

library(embed)

estimates <- 
  recipe(avg_price_per_room ~ ., data = hotel_tr) %>% 
  step_lencode_mixed(agent, outcome = vars(avg_price_per_room), id = "encoding") %>%   #<<
  step_dummy(all_nominal_predictors()) %>% 
  step_zv(all_predictors()) %>% 
  step_normalize(all_numeric_predictors()) %>% 
  prep() %>% 
  tidy(id = "encoding") %>% 
  select(agent = level, estimate = value)

before <- hotel_tr %>% select(avg_price_per_room, agent) %>% slice(1:7) %>% add_rowindex()
after <- left_join(before, estimates, by = "agent") %>% 
  select(avg_price_per_room, agent = estimate, .row)
```


## Per-player statistics 

::: columns
::: {.column width="50%"}
```{r}
#| label: effects-freq 
#| echo: false
#| out-width: '90%'
#| fig-width: 6
#| fig-height: 3
#| fig-align: 'center'
#| dev: 'svg'
#| dev-args: list(bg = "transparent")
  
agent_stats %>%   
  ggplot(aes(x = num_reservations)) +
  geom_histogram(bins = 30, col = "blue", fill = "blue", alpha = 1/3) +
  labs(x = "Number of reservations per player")
```
:::

::: {.column width="50%"}
```{r}
#| label: effects-adr 
#| echo: false
#| out-width: '90%'
#| fig-width: 6
#| fig-height: 3
#| fig-align: 'center'
#| dev: 'svg'
#| dev-args: list(bg = "transparent")

agent_stats %>%   
  ggplot(aes(x = ADR)) +
  geom_histogram(bins = 30, col = "red", fill = "red", alpha = 1/3) +
  labs(x = "Average ADR per agent")
```
:::
:::


## What is an effect encoding?

We replace the qualitative’s predictor data with their _effect on the outcome_. 

::: columns
::: {.column width="50%"}
Data before:

```{r}
#| label: before
before
```

:::

::: {.column width="50%"}

Data after:

```{r}
#| label: after
after
```

:::
:::

The `agent` column is replaced with an estimate of the ADR. 


## Per-agent statistics again {.annotation}

::: columns
::: {.column width="50%"}
```{r}
#| label: effects-again 
#| echo: false
#| out-width: '90%'
#| fig-width: 6
#| fig-height: 3
#| fig-align: 'center'
#| dev: 'svg'
#| dev-args: list(bg = "transparent")
  
agent_stats %>%   
  ggplot(aes(x = num_reservations)) +
  geom_histogram(bins = 30, col = "blue", fill = "blue", alpha = 1/3) +
  labs(x = "Number of reservations per agent")
agent_stats %>%   
  ggplot(aes(x = ADR)) +
  geom_histogram(bins = 30, col = "red", fill = "red", alpha = 1/3) +
  labs(x = "Average ADR per agent")
```
:::

::: {.column width="50%"}

- Good statistical methods for estimating these rates use *partial pooling*.


- Pooling borrows strength across players and shrinks extreme values (e.g. zero or one) towards the mean for players with very few shots.


- The embed package has recipe steps for effect encodings.

:::
:::


:::notes
Partial pooling gives better estimates for players with fewer shots by shrinking the estimate to the overall ADR mean


:::

## Partial pooling

```{r}
#| label: effect-compare 
#| echo: false
#| fig-width: 6
#| fig-height: 6
#| fig-align: 'center'
#| dev: 'svglite'
#| dev-args: list(bg = "transparent")

inner_join(agent_stats, estimates, by = "agent") %>% 
  ggplot(aes(x = ADR, y = estimate)) + 
  geom_abline(col = "green", lty = 2) +
  geom_point(aes(size = num_reservations), alpha = 1/3) +
  coord_obs_pred() +
  scale_size(range = c(1/3, 5)) +
  labs(x = "ADR Sample Mean", y = "Estimated via Effects Encoding")
```

## Player effects `r hexes("recipes","embed")` {.annotation}

```{r}
#| label: lencode-rec
#| code-line-numbers: "1,6|"
library(embed)

hotel_effect_rec <-
  recipe(avg_price_per_room ~ ., data = hotel_tr) %>% 
  step_YeoJohnson(lead_time) %>%
  step_lencode_mixed(agent, company, outcome = vars(avg_price_per_room)) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())
```

. . .

It is very important to appropriately validate the effect encoding step to make sure that we are not overfitting.

## Recipes are estimated `r hexes("recipes")`

Preprocessing steps in a recipe use the *training set* to compute quantities.

. . .

What kind of quantities are computed for preprocessing?

-   Levels of a factor
-   Whether a column has zero variance
-   Normalization
-   Feature extraction
-   Effect encodings

. . .

When a recipe is part of a workflow, this estimation occurs when `fit()` is called.

## Effect encoding results `r hexes("recipes","embed", "workflows", "tune")`

```{r}
#| label: resample-encoding
#| code-line-numbers: "4|"
hotel_effect_wflow <-
  workflow() %>%
  add_model(linear_reg()) %>% 
  update_recipe(hotel_effect_rec)

reg_metrics <- metric_set(mae, rsq)

hotel_effect_res <-
  hotel_effect_wflow %>%
  fit_resamples(hotel_rs, metrics = reg_metrics)

collect_metrics(hotel_effect_res)
```

Slightly worse but it can handle new agents (if they occur).

```{r}
#| label: teardown
#| include: false

parallel::stopCluster(cl)

```
