---
title: "6 - Feature selection"
subtitle: "Getting More Out of Feature Engineering and Tuning for Machine Learning"
format:
  revealjs: 
    slide-number: true
    footer: <https://workshops.tidymodels.org>
    include-before-body: header.html
    include-after-body: footer-annotations.html
    theme: [default, tidymodels.scss]
    width: 1280
    height: 720
knitr:
  opts_chunk: 
    echo: true
    collapse: true
    comment: "#>"
---

```{r setup}
#| include: false
#| file: setup.R
```


```{r}
#| label: start-quiet
#| include: false
library(mirai)
daemons(0) # reset to sequential if re-running
daemons(parallel::detectCores())
```

## Startup!  `r hexes(c("tidymodels"))`

```{r}
#| label: user-startup
library(tidymodels)
library(important)

tidymodels_prefer()
theme_set(theme_bw())
options(pillar.advice = FALSE, pillar.min_title_chars = Inf)
```

## More startup! 

```{r}
#| label: user-startup-2
# Load our example data for this section
"https://github.com/tidymodels/workshops/raw/refs/heads/2025-GMOFETML/slides/class_data.RData" |> 
  url() |> 
  load()

set.seed(429)
sim_split <- initial_split(class_data, prop = 0.75, strata = class)
sim_train <- training(sim_split)
sim_test  <- testing(sim_split)

set.seed(523)
sim_rs <- vfold_cv(sim_train, v = 10, strata = class)
```

# Why remove features/predictors?  {background-color="#E0D890FF"}

## Models and feature selection

Some models _automatically_ remove predictors by never using them in the model: 

 - tree- and rule-based models
 - some regularized models (e.g., `glmnet`)
 - multivariate adaptive regression splines
 - RuleFit
 - _not really_ ensembles though
 
Sometimes using irrelevant predictors *hurt* model performance

## Effects of extra predictors

![](images/fig-irrelevant-predictors-1.svg){fig-align='center' width=60%}

::: {.absolute bottom=0 right=0}
[AML4TD](https://aml4td.org/chapters/feature-selection.html#sec-irrelevant-predictors)
:::

## General selection methods

-  **wrappers**: a sequential algorithm proposes feature subsets, fits the model with these subsets, and then determines a better subset from the results

-  **filters**: screen predictors before adding them to the model. 

tidymodels doesn't have any wrappers (but see the [caret documentation for them](https://topepo.github.io/caret/recursive-feature-elimination.html))

<br> 

The new important package does have filters via recipes. 


## K-nearest neighbors

```{r}
#| label: knn-spec
rec <- 
  recipe(class ~ ., data = sim_train) |> 
  step_predictor_best(all_predictors(), score = "imp_rf", prop_terms = tune()) |> 
  step_normalize(all_numeric_predictors())
  
knn_spec <- 
  nearest_neighbor(neighbors = tune(), weight_func = tune()) |> 
  set_mode("classification")
  
thrsh_tlr <-
  tailor() |>
  adjust_probability_threshold(threshold = tune()) 
  
knn_wflow <- workflow(rec, knn_spec, thrsh_tlr)

knn_param <- 
  knn_wflow |> 
  extract_parameter_set_dials()
```

## Tuning results

```{r}
#| label: knn-tune
#| cache: true
cls_mtr <- metric_set(brier_class, roc_auc, sensitivity, specificity)
ctrl <- control_grid(save_pred = TRUE, save_workflow = TRUE)

set.seed(12)
knn_res <-
  knn_wflow |>
  tune_grid(
    resamples = sim_rs,
    grid = 50,
    control = ctrl,
    metrics = cls_mtr
  )
```

## Grid results `r hexes(c("tune"))`

```{r}
#| label: sim-knn-all
#| fig-width: 12
#| fig-height: 5
#| fig-align: center
#| out-width: "100%"
#| dev-args:
#|   bg: "transparent"
autoplot(knn_res)
```

## Brier results `r hexes(c("tune"))`

```{r}
#| label: sim-knn-brier
#| fig-width: 12
#| fig-height: 3
#| fig-align: center
#| out-width: "100%"
#| dev-args:
#|   bg: "transparent"
autoplot(knn_res, metric = "brier_class") + 
  facet_grid(. ~ name, scale = "free_x") 
```

## ROC curve results `r hexes(c("tune"))`

```{r}
#| label: sim-knn-roc
#| fig-width: 12
#| fig-height: 3
#| fig-align: center
#| out-width: "100%"
#| dev-args:
#|   bg: "transparent"
autoplot(knn_res, metric = "roc_auc") + 
  facet_grid(. ~ name, scale = "free_x") 
```

## Sensitivity/Specificity results `r hexes(c("tune"))`

```{r}
#| label: sim-knn-two-class
#| fig-width: 12
#| fig-height: 4.5
#| fig-align: center
#| out-width: "100%"
#| dev-args:
#|   bg: "transparent"
autoplot(knn_res, metric = c("sensitivity", "specificity"))
```



