---
title: "Extras - Time-based data splitting"
subtitle: "Introduction to tidymodels"
format:
  revealjs: 
    slide-number: true
    footer: <https://workshops.tidymodels.org>
    include-before-body: header.html
    include-after-body: footer-annotations.html
    theme: [default, tidymodels.scss]
    width: 1280
    height: 720
knitr:
  opts_chunk: 
    echo: true
    collapse: true
    comment: "#>"
    fig.path: "figures/"
---

```{r}
#| label: setup
#| include: false
#| file: setup.R
```

```{r}
#| label: setup-previous
#| echo: false
library(countdown)
library(tidymodels)
library(modeldatatoo)

taxi_raw <- readr::read_rds("taxi_raw.rds")
```

## The raw taxi data set

We prepared the data set specifically for this introductory workshop. 

It looked similar to this:

```{r}
glimpse(taxi_raw)
```

:::notes
`trip_start_time` has date and time rounded to "quarters of the hour".
::::


## Time nature of the data

We assumed only the month, day of the week, and hour mattered and treated each observation as independent.

. . . 

If the data have a strong time component, all your data splitting strategies should support the model in estimating temporal trends.

. . .

Thus, don't sample randomly because this breaks up the time component!


## Splitting with time component `r hexes(c("rsample"))`

The more recent observations are assumed to be more similar to new data, so `initial_time_split()` puts them into the test set. 

The function assumes that the data are already ordered.

```{r}
#| label: split

taxi_raw <- taxi_raw %>%
  arrange(trip_start_timestamp)

taxi_split <- initial_time_split(taxi_raw, prop = 3 / 4)
taxi_split

taxi_train <- training(taxi_split)
taxi_test  <- testing(taxi_split)

nrow(taxi_train)
 
nrow(taxi_test)
```

## Time series resampling 

The same idea also applies to resampling: the newer observations go into the assessment set.

For example:

* Fold 1: Take the first X weeks of data as the analysis set, and the next 2 weeks as the assessment set.

* Fold 2: Take the first X weeks + 2 weeks of data as the analysis set, and the next 2 weeks as the assessment set.

* and so on

## Rolling origin forecast resampling 

```{r}
#| label: rolling
#| echo: false
#| out.width: 65%
#| fig.align: center
#| out-width: "70%"

knitr::include_graphics("images/rolling.svg")
```

:::notes
This image shows overlapping assessment sets. We will use non-overlapping data but it could be done either way.
:::

## Times series resampling  `r hexes(c("rsample"))`

```{r}
#| eval: false
#| code-line-numbers: "4"

taxi_rs <-
  taxi_train %>%
  sliding_period(
    index = "trip_start_timestamp",  




  )
```

Use the `trip_start_timestamp` column to find the date data. 

## Times series resampling  `r hexes(c("rsample"))`

```{r}
#| eval: false
#| code-line-numbers: "5"

taxi_rs <-
  taxi_train %>%
  sliding_period(
    index = "trip_start_timestamp",  
    period = "week",



  )
```

Our units will be in weeks. 

## Times series resampling  `r hexes(c("rsample"))`

```{r}
#| eval: false
#| code-line-numbers: "6"

taxi_rs <-
  taxi_train %>%
  sliding_period(
    index = "trip_start_timestamp",  
    period = "week",
    lookback = 5
    
    
  )
```

Every analysis set has 5 weeks of data.

## Times series resampling  `r hexes(c("rsample"))`

```{r}
#| eval: false
#| code-line-numbers: "7"

taxi_rs <-
  taxi_train %>%
  sliding_period(
    index = "trip_start_timestamp",  
    period = "week",
    lookback = 5,
    assess_stop = 2,

  )
```

Every assessment set has 2 weeks of data.

## Times series resampling  `r hexes(c("rsample"))`

```{r}
#| code-line-numbers: "8"

taxi_rs <-
  taxi_train %>%
  sliding_period(
    index = "trip_start_timestamp",  
    period = "week",
    lookback = 5,
    assess_stop = 2,
    step = 2 
  )
```

Increment by 2 weeks so that there are no overlapping assessment sets. 

. . .

```{r}
taxi_rs$splits[[1]] %>% assessment() %>% pluck("trip_start_timestamp") %>% range()

taxi_rs$splits[[2]] %>% assessment() %>% pluck("trip_start_timestamp") %>% range()

taxi_rs$splits[[3]] %>% assessment() %>% pluck("trip_start_timestamp") %>% range()
```
