---
title: "6 - Tuning Hyperparameters - Classwork"
subtitle: "Machine learning with tidymodels"
editor_options: 
  chunk_output_type: console
---

We recommend restarting R between each slide deck!

## Setup

Setup from deck 5

```{r}
library(tidymodels)
library(embed)
library(ongoal)

tidymodels_prefer()

set.seed(23)
nhl_split <- initial_split(season_2015, prop = 3/4)
nhl_split

nhl_train_and_val <- training(nhl_split)
nhl_test  <- testing(nhl_split)

set.seed(234)
nhl_val <- validation_split(nhl_train_and_val, prop = 0.80)

nhl_train <- analysis(nhl_val$splits[[1]])

nhl_position_rec <-
  recipe(on_goal ~ ., data = nhl_train) %>%
  step_lencode_mixed(shooter, goaltender, outcome = vars(on_goal)) %>%
  step_other(all_nominal_predictors()) %>% # TODO: keep this?
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors()) %>%
  step_mutate(
    angle = abs( atan2(abs(coord_y), (89 - coord_x) ) * (180 / pi)),
    behind_goal_line = ifelse(coord_x >= 89, 1, 0)
  )

nhl_position_wflow <-
  workflow() %>%
  add_recipe(nhl_position_rec) %>%
  add_model(logistic_reg())

nhl_position_res <-
  nhl_position_wflow %>%
  fit_resamples(nhl_val)
```

## Updates for tuning

```{r}
glm_rec <-
  recipe(on_goal ~ ., data = nhl_train) %>%
  step_lencode_mixed(shooter, goaltender, outcome = vars(on_goal)) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_mutate(
    angle = abs( atan2(abs(coord_y), (89 - coord_x) ) * (180 / pi) ),
    defensive_zone = ifelse(coord_x <= -25.5, 1, 0),
    behind_goal_line = ifelse(coord_x >= 89, 1, 0)
  ) %>%
  step_zv(all_predictors()) %>%
  step_ns(angle, deg_free = tune("angle")) %>%
  step_ns(coord_x, deg_free = tune("coord_x")) %>%
  step_normalize(all_numeric_predictors())

glm_spline_wflow <-
  workflow() %>%
  add_model(logistic_reg()) %>%
  add_recipe(glm_rec)
```

## Create a grid

```{r}
set.seed(12)
grid <- 
  glm_spline_wflow %>% 
  extract_parameter_set_dials() %>% 
  grid_latin_hypercube(size = 25)
```

## Your turn

Create a grid for our tunable workflow.

Try creating a regular grid.

```{r}
# Your code here!

```

## Update parameter ranges

```{r}
set.seed(12)
grid <- 
  glm_spline_wflow %>% 
  extract_parameter_set_dials() %>% 
  update(angle = spline_degree(c(2L, 50L)),
         coord_x = spline_degree(c(2L, 50L))) %>% 
  grid_latin_hypercube(size = 25)

grid %>% 
  ggplot(aes(angle, coord_x)) +
  geom_point(size = 4)
```

## Spline grid search

```{r} 
set.seed(9)
ctrl <- control_grid(save_pred = TRUE, parallel_over = "everything")

glm_spline_res <-
  glm_spline_wflow %>%
  tune_grid(resamples = nhl_val, grid = grid, control = ctrl)

glm_spline_res
```

## Your turn

Tune our `glm_wflow`.

What happens if you don't supply a `grid` argument to `tune_grid()`?

```{r}
# Your code here!

```

## Grid results

```{r}
autoplot(glm_spline_res)
```

## Tuning results

```{r}
collect_metrics(glm_spline_res)
collect_metrics(glm_spline_res, summarize = FALSE)
```

## Choose a parameter combination

```{r}
show_best(glm_spline_res, metric = "roc_auc")
select_best(glm_spline_res, metric = "roc_auc")
```

## Boosted trees

```{r}
xgb_spec <-
  boost_tree(
    trees = tune(), min_n = tune(), tree_depth = tune(),
    learn_rate = tune(), loss_reduction = tune()
  ) %>%
  set_mode("classification") %>% 
  set_engine("xgboost") 

xgb_rec <- 
  recipe(on_goal ~ ., data = nhl_train) %>% 
  step_lencode_mixed(shooter, goaltender, outcome = vars(on_goal)) %>% 
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())

xgb_wflow <- 
  workflow() %>% 
  add_model(xgb_spec) %>% 
  add_recipe(xgb_rec)
```

## Your turn

Create your boosted tree workflow.

```{r}
# Your code here!

```

## Running in parallel

```{r}
cores <- parallelly::availableCores(logical = FALSE)
cl <- parallel::makePSOCKcluster(cores)
doParallel::registerDoParallel(cl)
```

## Tuning

```{r}
# this will take some time to run
set.seed(9)

xgb_res <-
  xgb_wflow %>%
  tune_grid(resamples = nhl_val, grid = 30, control = ctrl) # automatic grid now!
```

## Your turn 

Start tuning the boosted tree model!

We won't wait for everyone's tuning to finish, but take this time to get it started before we move on.

```{r}
# Your code here!

```

## Tuning results

```{r}
xgb_res

autoplot(xgb_res)
```

## Again with the location features

```{r}
coord_rec <- 
  xgb_rec %>%
  step_mutate(
    angle = abs( atan2(abs(coord_y), (89 - coord_x) ) * (180 / pi) ),
    defensive_zone = ifelse(coord_x <= -25.5, 1, 0),
    behind_goal_line = ifelse(coord_x >= 89, 1, 0)
  )

xgb_coord_wflow <- 
  workflow() %>% 
  add_model(xgb_spec) %>% 
  add_recipe(coord_rec)

set.seed(9)
xgb_coord_res <-
  xgb_coord_wflow %>%
  tune_grid(resamples = nhl_val, grid = 30, control = ctrl)
```

## Did the machine figure it out? 

```{r}
# no extra features
show_best(xgb_res, metric = "roc_auc", n = 3)
# with additional coordinate features
show_best(xgb_coord_res, metric = "roc_auc", n = 3)
```

## Compare models

```{r}
# Best logistic regression results
glm_spline_res %>% 
  show_best(metric = "roc_auc", n = 1) %>% 
  select(.metric, .estimator, mean, n, std_err, .config)
```

```{r}
# Best boosting results
xgb_res %>% 
  show_best(metric = "roc_auc", n = 1) %>% 
  select(.metric, .estimator, mean, n, std_err, .config)
```

## Updating the workflow

```{r}
best_auc <- select_best(glm_spline_res, metric = "roc_auc")
best_auc

glm_spline_wflow <-
  glm_spline_wflow %>% 
  finalize_workflow(best_auc)

glm_spline_wflow
```

## The final fit 

```{r}
test_res <- 
  glm_spline_wflow %>% 
  last_fit(split = nhl_split)

test_res
```

## Your turn 

Finalize your workflow with the best parameters.

Create a final fit.

```{r}
# Your code here!

```

## Estimates of ROC AUC

```{r}
# Validation results from tuning
glm_spline_res %>% 
  show_best(metric = "roc_auc", n = 1) %>% 
  select(.metric, mean, n, std_err)

# Test set results
test_res %>% collect_metrics()
```

## Final fitted workflow

```{r}
final_glm_spline_wflow <- 
  test_res %>% 
  extract_workflow()

# use this object to predict or deploy
predict(final_glm_spline_wflow, nhl_test[1:3,])
```

## Explain yourself

Create an explainer for our glm model.

```{r}
library(DALEXtra)

glm_explainer <- explain_tidymodels(
  final_glm_spline_wflow,
  data = dplyr::select(nhl_train, -on_goal),
  # DALEX required an integer for factors:
  y = as.integer(nhl_train$on_goal) - 1,
  verbose = FALSE
)
```

Create partial dependence profiles

https://ema.drwhy.ai/partialDependenceProfiles.html

```{r}
set.seed(123)
pdp_coord_x <- model_profile(
  glm_explainer,
  variables = "coord_x",
  N = 500,
  groups = "strength"
)
```

## Your turn 

Try grouping by another variable, like `extra_attacker` or `game_seconds`.

```{r}
# Your code here!

```


```{r}
# turn off parallel backend
foreach::registerDoSEQ()
parallel::stopCluster(cl)
```
