---
title: "2 - Your data budget - Classwork"
subtitle: "Practical Machine Learning with tidymodels"
editor_options:
  chunk_output_type: console
---

We recommend restarting R between each slide deck!

## A classification data set

```{r}
library(tidymodels)
# cls_data_2026 is already loaded from the modeldata package
glimpse(cls_data_2026)
```

## Your turn

Let's take 5 minutes and look at the data.

```{r}
# Your code here!

```

## Your turn

When is a good time to split your data?

## The initial split

```{r}
set.seed(123)
cls_split <- initial_split(cls_data_2026)
cls_split
```

## Accessing the data

```{r}
cls_train <- training(cls_split)
cls_test <- testing(cls_split)
```

## The training set

```{r}
cls_train
```

## Your turn

Split your data so 20% is held out for the test set.

Try out different values in `set.seed()` to see how the results change.

Hint: Which argument in `initial_split()` handles the proportion split into
training vs testing?

```{r}
# Your code here!

```

## Data splitting and spending

```{r}
set.seed(123)
cls_split <- initial_split(cls_data_2026, prop = 0.8)
cls_train <- training(cls_split)
cls_test <- testing(cls_split)

nrow(cls_train)
nrow(cls_test)
```

## Your turn

Explore the `cls_train` data on your own!

- What's the distribution of the outcome, `class`?
- What's the distribution of numeric variables?

```{r}
# Your code here!

```
