---
title: "2 - Your data budget - Classwork"
subtitle: "Machine learning with tidymodels"
editor_options: 
  chunk_output_type: console
---

We recommend restarting R between each slide deck!

## Data on tree frog hatching

```{r}
library(tidymodels)

data("tree_frogs", package = "stacks")

# Slightly modify the original data for the purposes of this workshop
tree_frogs <- tree_frogs %>%
  mutate(t_o_d = factor(t_o_d),
         age = age / 86400) %>%
  filter(!is.na(latency)) %>%
  select(-c(clutch, hatched))
```

```{r}
tree_frogs
```

## Your turn

When is a good time to split your data?

## Data splitting and spending

```{r}
set.seed(123)

# Split into training/testing, stratifying by `latency`
frog_split <- initial_split(tree_frogs, strata = latency)
frog_split
```

Extract the training and testing sets

```{r}
frog_train <- training(frog_split)
frog_test <- testing(frog_split)
```

## Your turn

Split your data so 20% is held out for the test set.

Try out different values in `set.seed()` to see how the results change.

Hint: Which argument in `initial_split()` handles the proportion split into training vs testing?

```{r}
# Your code here!

```

## Your turn

Explore the `frog_train` data on your own!

- What's the distribution of the outcome, latency?
- What's the distribution of numeric variables like age?
- How does latency differ across the categorical variables?

```{r}
# Your code here!

```

## Stratification

```{r}
set.seed(123)

frog_split <- initial_split(tree_frogs, prop = 0.8, strata = latency)
frog_train <- training(frog_split)
frog_test <- testing(frog_split)
```
