---
title: "5 - Feature engineering - Classwork"
subtitle: "Machine learning with tidymodels"
editor_options: 
  chunk_output_type: console
---

We recommend restarting R between each slide deck!

## Case study

```{r}
library(tidymodels)
library(ongoal)

tidymodels_prefer()
ggplot2::theme_set(ggplot2::theme_bw())

glimpse(season_2015)
```

## Splitting the NHL data

```{r}
set.seed(23)
nhl_split <- initial_split(season_2015, prop = 3/4)
nhl_split

nhl_train <- training(nhl_split)
nhl_test  <- testing(nhl_split)
```

## Your turn

Let's explore the training set data.

Use the function `plot_nhl_shots()` for nice spatial plots of the data.

```{r}
set.seed(100)
nhl_train %>% 
  sample_n(200) %>%
  plot_nhl_shots(emphasis = position)

# Your code here!

```

## Validation split

```{r}
set.seed(234)
nhl_val <- validation_split(nhl_train, prop = 0.80)
nhl_val
```

## A first recipe

```{r}
nhl_rec <- 
  recipe(on_goal ~ ., data = nhl_train)
summary(nhl_rec)
```

## A basic recipe

```{r}
nhl_rec <- 
  recipe(on_goal ~ ., data = nhl_train) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_zv(all_predictors()) %>% 
  step_normalize(all_numeric_predictors())
```

## Other possible steps

```{r}
# Reduce correlation 
nhl_rec <- 
  recipe(on_goal ~ ., data = nhl_train) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_zv(all_predictors()) %>% 
  step_normalize(all_numeric_predictors()) %>% 
  step_corr(all_numeric_predictors(), threshold = 0.9)

# PCA feature extraction
nhl_rec <- 
  recipe(on_goal ~ ., data = nhl_train) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_zv(all_predictors()) %>% 
  step_normalize(all_numeric_predictors()) %>% 
  step_pca(all_numeric_predictors())

# Supervised and unsupervised uniform manifold approximation and projection (UMAP)
library(embed)
nhl_rec <- 
  recipe(on_goal ~ ., data = nhl_train) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_zv(all_predictors()) %>% 
  step_normalize(all_numeric_predictors()) %>% 
  embed::step_umap(all_numeric_predictors(), outcome = on_goal)

# Natural splines
nhl_rec <- 
  recipe(on_goal ~ ., data = nhl_train) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_zv(all_predictors()) %>% 
  step_normalize(all_numeric_predictors()) %>% 
  step_ns(coord_y, coord_x, deg_free = 10)
```

## Your turn

Create a `recipe()` for the on-goal data to :

-   create dummy variables
-   remove zero-variance variables

```{r}
# Your code here!

```

## Minimal recipe 

```{r}
nhl_indicators <-
  recipe(on_goal ~ ., data = nhl_train) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())
```

## Using a workflow

```{r}
nhl_glm_wflow <-
  workflow() %>%
  add_recipe(nhl_indicators) %>%
  add_model(logistic_reg())
 
ctrl <- control_resamples(save_pred = TRUE)
nhl_glm_res <-
  nhl_glm_wflow %>%
  fit_resamples(nhl_val, control = ctrl)

collect_metrics(nhl_glm_res)
```

## Your turn

Use `fit_resamples()` to fit your workflow with a recipe.

Collect the predictions from the results.

```{r}
# Your code here!

```

## Holdout predictions

```{r}
# Since we used `save_pred = TRUE`
glm_val_pred <- collect_predictions(nhl_glm_res)
glm_val_pred %>% slice(1:7)
```

## ROC curves

```{r}
# Assumes _first_ factor level is event; there are options to change that
roc_curve_points <- glm_val_pred %>% roc_curve(truth = on_goal, estimate = .pred_yes)
roc_curve_points %>% slice(1, 50, 100)

glm_val_pred %>% roc_auc(truth = on_goal, estimate = .pred_yes)
```

## ROC curve plot 

```{r}
autoplot(roc_curve_points)
```

## Your turn

Compute and plot an ROC curve for your current model.

```{r}
# Your code here!

```

## Your turn

What data is being used for this ROC curve plot?

## Player effects

```{r}
library(embed)

nhl_effect_rec <-
  recipe(on_goal ~ ., data = nhl_train) %>%
  step_lencode_mixed(player, outcome = vars(on_goal)) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())
```

## Effect encoding results

```{r}
nhl_effect_wflow <-
  nhl_glm_wflow %>%
  update_recipe(nhl_effect_rec)

nhl_effect_res <-
  nhl_effect_wflow %>%
  fit_resamples(nhl_val)

collect_metrics(nhl_effect_res)
```

## Where is the shot coming from?

```{r}
# angle
nhl_angle_rec <-
  nhl_indicators %>%
  step_mutate(
    angle = abs(atan2(abs(coord_y), (89 - abs(coord_x))) * (180 / pi))
  )

# distance
nhl_distance_rec <-
  nhl_angle_rec %>%
  step_mutate(
    distance = sqrt((89 - abs(coord_x))^2 + abs(coord_y)^2),
    distance = log(distance)
  )

# behind goal line
nhl_behind_rec <-
  nhl_distance_rec %>%
  step_mutate(
    behind_goal_line = ifelse(abs(coord_x) >= 89, 1, 0)
  )
```

## Fit different recipes

```{r}
nhl_glm_set_res <-
  workflow_set(
    list(dummy = nhl_indicators, encoded = nhl_effect_rec,
         angle = nhl_angle_rec, dist = nhl_distance_rec, 
         bgl = nhl_behind_rec),
    list(logistic = logistic_reg())
  ) %>%
  workflow_map(fn = "fit_resamples", resamples = nhl_val, verbose = TRUE, control = ctrl)
```

## Your turn

Create a workflow set with 2 or 3 recipes.

Use `workflow_map()` to resample the workflow set.

```{r}
# Your code here!

```

## Compare recipes

```{r}
library(forcats)

collect_metrics(nhl_glm_set_res) %>%
  filter(.metric == "roc_auc") %>%
  mutate(
    features = gsub("_logistic", "", wflow_id), 
    features = fct_reorder(features, mean)
  ) %>%
  ggplot(aes(x = mean, y = features)) +
  geom_point(size = 3) +
  labs(y = NULL, x = "ROC AUC (validation set)")
```
