---
title: "1 - Feature Engineering - Classwork"
subtitle: "Advanced tidymodels"
editor_options: 
  chunk_output_type: console
---

We recommend restarting R between each slide deck!

## Hotel data

```{r}
library(tidymodels)
library(modeldatatoo)

# Max's usual settings: 
tidymodels_prefer()
theme_set(theme_bw())
options(
  pillar.advice = FALSE, 
  pillar.min_title_chars = Inf
)

set.seed(295)
hotel_rates <- 
  data_hotel_rates() %>% 
  sample_n(5000) %>% 
  arrange(arrival_date) %>% 
  select(-arrival_date_num, -arrival_date) %>% 
  mutate(
    company = factor(as.character(company)),
    country = factor(as.character(country)),
    agent = factor(as.character(agent))
  )
```

## Data spending

```{r}
set.seed(4028)
hotel_split <-
  initial_split(hotel_rates, strata = avg_price_per_room)

hotel_tr <- training(hotel_split)
hotel_te <- testing(hotel_split)
```

## Your turn

Let's take some time and investigate the _training data_. The outcome is `avg_price_per_room`. 

Are there any interesting characteristics of the data?

```{r}
# Your code here!

```

## Resampling Strategy

```{r}
set.seed(472)
hotel_rs <- vfold_cv(hotel_tr, strata = avg_price_per_room)
hotel_rs
```

## A first recipe

```{r}
hotel_rec <- 
  recipe(avg_price_per_room ~ ., data = hotel_tr)

summary(hotel_rec)
```

## Your turn

What do you think are in the `type` vectors for the `lead_time` and `country` columns?

```{r}
# Your code here!

```

## A base recipe

```{r}
hotel_rec <- 
  recipe(avg_price_per_room ~ ., data = hotel_tr) %>% 
  # create indicator variables
  step_dummy(all_nominal_predictors()) %>% 
  # filter out constant columns
  step_zv(all_predictors()) %>% 
  # normalize
  step_normalize(all_numeric_predictors())
```

## Different options to reduce correlation

```{r}
hotel_rec <- 
  recipe(avg_price_per_room ~ ., data = hotel_tr) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_zv(all_predictors()) %>% 
  step_normalize(all_numeric_predictors()) %>% 
  step_corr(all_numeric_predictors(), threshold = 0.9)

hotel_rec <- 
  recipe(avg_price_per_room ~ ., data = hotel_tr) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_zv(all_predictors()) %>% 
  step_normalize(all_numeric_predictors()) %>% 
  step_pca(all_numeric_predictors())

hotel_rec <- 
  recipe(avg_price_per_room ~ ., data = hotel_tr) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_zv(all_predictors()) %>% 
  step_normalize(all_numeric_predictors()) %>% 
  embed::step_umap(all_numeric_predictors(), outcome = avg_price_per_room)
```

## Other possible steps

For example, natural splines:

```{r}
hotel_rec <- 
  recipe(avg_price_per_room ~ ., data = hotel_tr) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_zv(all_predictors()) %>% 
  step_normalize(all_numeric_predictors()) %>% 
  step_spline_natural(year_day, deg_free = 10)
```

## Your turn 

Create a `recipe()` for the hotel data to:

- use a Yeo-Johnson (YJ) transformation on `lead_time`
- convert factors to indicator variables
- remove zero-variance variables

```{r}
# Your code here!

```

## Minimal recipe for the hotel data

```{r}
hotel_indicators <-
  recipe(avg_price_per_room ~ ., data = hotel_tr) %>% 
  step_YeoJohnson(lead_time) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())
```


## Measuring Performance

We'll compute two measures, mean absolute error (MAE) and the coefficient of determination (a.k.a $R^2$), and focus on the MAE for parameter optimization. 

```{r}
reg_metrics <- metric_set(mae, rsq)
```

```{r}
set.seed(9)

hotel_lm_wflow <-
  workflow() %>%
  add_recipe(hotel_indicators) %>%
  add_model(linear_reg())
 
ctrl <- control_resamples(save_pred = TRUE)
hotel_lm_res <-
  hotel_lm_wflow %>%
  fit_resamples(hotel_rs, control = ctrl, metrics = reg_metrics)

collect_metrics(hotel_lm_res)
```

## Your turn

Use `fit_resamples()` to fit your workflow with a recipe.

Collect the predictions from the results.

```{r}
# Your code here!

```

## Holdout predictions

```{r}
# Since we used `save_pred = TRUE`
lm_val_pred <- collect_predictions(hotel_lm_res)
lm_val_pred %>% slice(1:7)
```

## Calibration Plot 

```{r}
library(probably)

cal_plot_regression(hotel_lm_res, alpha = 1 / 5)
```

## What do we do with the agent and company data? 

Collapsing factor levels: 

```{r}
hotel_other_rec <-
  recipe(avg_price_per_room ~ ., data = hotel_tr) %>% 
  step_YeoJohnson(lead_time) %>%
  step_other(agent, threshold = 0.001) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())

hotel_other_wflow <-
  hotel_lm_wflow %>%
  update_recipe(hotel_other_rec)

hotel_other_res <-
  hotel_other_wflow %>%
  fit_resamples(hotel_rs, control = ctrl, metrics = reg_metrics)

collect_metrics(hotel_other_res)
```

Feature Hashing:

```{r}
library(textrecipes)

hash_rec <-
  recipe(avg_price_per_room ~ ., data = hotel_tr) %>%
  step_YeoJohnson(lead_time) %>%
  # Defaults to 32 signed indicator columns
  step_dummy_hash(agent) %>%
  step_dummy_hash(company) %>%
  # Regular indicators for the others
  step_dummy(all_nominal_predictors()) %>% 
  step_zv(all_predictors())

hotel_hash_wflow <-
  hotel_lm_wflow %>%
  update_recipe(hash_rec)

hotel_hash_res <-
  hotel_hash_wflow %>%
  fit_resamples(hotel_rs, control = ctrl, metrics = reg_metrics)

collect_metrics(hotel_hash_res)
```

## Debugging a recipe

```{r}
# Estimate the transformation coefficients
hash_rec_fit <- prep(hash_rec)

# Get the transformation coefficient
tidy(hash_rec_fit, number = 1)

# Get the processed data
bake(hash_rec_fit, hotel_tr %>% slice(1:3), contains("_agent_"))
```
