---
title: "2 - Tuning Hyperparameters - Classwork"
subtitle: "Advanced tidymodels"
editor_options: 
  chunk_output_type: console
---

We recommend restarting R between each slide deck!

## Setup

Setup from deck 2

```{r}
library(tidymodels)
library(textrecipes)

# Max's usual settings: 
tidymodels_prefer()
theme_set(theme_bw())
options(
  pillar.advice = FALSE, 
  pillar.min_title_chars = Inf
)

reg_metrics <- metric_set(mae, rsq)

set.seed(295)
hotel_rates <- 
  hotel_rates %>% 
  sample_n(5000) %>% 
  arrange(arrival_date) %>% 
  select(-arrival_date) %>% 
  mutate(
    company = factor(as.character(company)),
    country = factor(as.character(country)),
    agent = factor(as.character(agent))
  )

set.seed(4028)
hotel_split <-
  initial_split(hotel_rates, strata = avg_price_per_room)

hotel_train <- training(hotel_split)
hotel_test <- testing(hotel_split)

set.seed(472)
hotel_rs <- vfold_cv(hotel_train, strata = avg_price_per_room)

hash_rec <-
  recipe(avg_price_per_room ~ ., data = hotel_train) %>%
  step_YeoJohnson(lead_time) %>%
  # Defaults to 32 signed indicator columns
  step_dummy_hash(agent) %>%
  step_dummy_hash(company) %>%
  # Regular indicators for the others
  step_dummy(all_nominal_predictors()) %>% 
  step_zv(all_predictors())
```

## Tagging parameters for tuning

```{r}
hash_rec <-
  recipe(avg_price_per_room ~ ., data = hotel_train) %>%
  step_YeoJohnson(lead_time) %>%
  step_dummy_hash(agent,   num_terms = tune("agent hash")) %>%
  step_dummy_hash(company, num_terms = tune("company hash")) %>%
  step_zv(all_predictors())
```

```{r}
library(bonsai)

lgbm_spec <- 
  boost_tree(trees = tune(), learn_rate = tune()) %>% 
  set_mode("regression") %>% 
  set_engine("lightgbm")

lgbm_wflow <- workflow(hash_rec, lgbm_spec)
```

## Create a grid 

```{r}
set.seed(12)
grid <- 
  lgbm_wflow %>% 
  extract_parameter_set_dials() %>% 
  grid_latin_hypercube(size = 25)

grid
```

## Your turn 

Create a grid for our tunable workflow.

Try creating a regular grid.

```{r}
# Your code here!

```

## Your turn 

What advantage would a regular grid have? 

Discuss with your neighbor!

## Update parameter ranges 

```{r} 
lgbm_param <- 
  lgbm_wflow %>% 
  extract_parameter_set_dials() %>% 
  update(trees = trees(c(1L, 100L)),
         learn_rate = learn_rate(c(-5, -1)))

set.seed(712)
grid <- 
  lgbm_param %>% 
  grid_latin_hypercube(size = 25)

grid
```

```{r}
grid %>% 
  ggplot(aes(trees, learn_rate)) +
  geom_point(size = 4) +
  scale_y_log10()
```

## Grid Search

Let's take our previous model and tune more parameters:

```{r} 
lgbm_spec <- 
  boost_tree(trees = tune(), learn_rate = tune(),  min_n = tune()) %>% 
  set_mode("regression") %>% 
  set_engine("lightgbm")

lgbm_wflow <- workflow(hash_rec, lgbm_spec)

# Update the feature hash ranges (log-2 units)
lgbm_param <-
  lgbm_wflow %>%
  extract_parameter_set_dials() %>%
  update(`agent hash`   = num_hash(c(3, 8)),
         `company hash` = num_hash(c(3, 8)))
```

Run the grid search:

```{r} 
set.seed(9)
ctrl <- control_grid(save_pred = TRUE)

lgbm_res <-
  lgbm_wflow %>%
  tune_grid(
    resamples = hotel_rs,
    grid = 25,
    # The options below are not required by default
    param_info = lgbm_param, 
    control = ctrl,
    metrics = reg_metrics
  )

lgbm_res 
```

Inspect results:

```{r}
autoplot(lgbm_res)

collect_metrics(lgbm_res)
collect_metrics(lgbm_res, summarize = FALSE)
```

## Choose a parameter combination

```{r}
show_best(lgbm_res, metric = "rsq")

lgbm_best <- select_best(lgbm_res, metric = "mae")
lgbm_best
```

## Checking Calibration

```{r}
library(probably)

lgbm_res %>%
  collect_predictions(
    parameters = lgbm_best
  ) %>%
  cal_plot_regression(
    truth = avg_price_per_room,
    estimate = .pred,
    alpha = 1 / 3
  )
```

## Running in parallel

```{r}
#| eval: false

cores <- parallelly::availableCores(logical = FALSE)
cl <- parallel::makePSOCKcluster(cores)
doParallel::registerDoParallel(cl)

# Now call `tune_grid()`!

# Shut it down with:
foreach::registerDoSEQ()
parallel::stopCluster(cl)
```

## Your turn

Try early stopping: Set `trees = 2000` and tune the `stop_iter` parameter!

Note that you will need to regenerate `lgbm_param` with your new workflow!

```{r}
# Your code here!

```
