---
title: "7 - Feature engineering: splines, target encoding and dates - Classwork"
subtitle: "Practical Machine Learning with tidymodels"
editor_options: 
  chunk_output_type: console
---

We recommend restarting R between each slide deck!

## Setup

```{r}
library(tidymodels)
library(embed)
library(extrasteps)

tidymodels_prefer()
theme_set(theme_bw())
options(pillar.advice = FALSE, pillar.min_title_chars = Inf)
```

## Hotel rates data set

```{r}
glimpse(hotel_rates)
```

## Hotel data splitting

```{r}
set.seed(1234)
hotel_split <- initial_split(hotel_rates)
hotel_train <- training(hotel_split)
hotel_test <- testing(hotel_split)
```

## Your turn

Load and explore the `hotel_train` data.

Comes loaded with the modeldata package.

```{r}
# Your code here!

```

## Your turn

Apply B-splines to some variables using `step_spline_b()`.

```{r}
# Your code here!

```

## Factors with many categories

```{r}
hotel_train |>
  count(country)
```

```{r}
hotel_train |>
  count(company)
```

```{r}
hotel_train |>
  count(agent)
```

## Target encoding motivation

We calculate the mean of the outcome within each level of the predictor, and use that as the new value.

Don't do just this! We are building up the method one thing at a time. Unregularized target encoding is really prone to overfitting.

```{r}
hotel_train |>
  summarise(
    mean = mean(avg_price_per_room),
    .by = agent
  )
```

## Target encoding handling unseen levels

```{r}
hotel_train |>
  count(agent, .drop = FALSE)
```

```{r}
hotel_test |>
  count(agent, .drop = FALSE)
```

Calculate the global mean of the outcome and use it for cases that aren't seen in the training data set.

```{r}
mean(hotel_train$avg_price_per_room)
```

```{r}
hotel_train |>
  summarise(
    mean = mean(avg_price_per_room),
    .by = agent
  )
```

## How do we handle low counts?

```{r}
hotel_train |>
  summarise(
    mean = mean(avg_price_per_room),
    n = n(),
    .by = agent
  ) |>
  arrange(agent)
```

## Target encoding in recipes

```{r}
recipe(avg_price_per_room ~ ., data = hotel_train) |>
  step_lencode(
    agent,
    country,
    company,
    outcome = vars("avg_price_per_room"),
    smooth = TRUE,
  ) |>
  prep() |>
  bake(NULL) |>
  select(agent, country, company)
```

## Your turn

Apply target encoding to the data set, see how it affects different predictors, not just the ones we listed here.

* `step_lencode()`
* `step_lencode_glm()`
* `step_lencode_bayes()`
* `step_lencode_mixed()`

```{r}
# Your code here!

```

## Your turn

Explore the `arrival_date` variable and its relation to `avg_price_per_room`.

The lubridate package might be helpful.

```{r}
# Your code here!

```

## Your turn

Apply date steps to the `arrival_date` variable and try to see if we capture anything about `avg_price_per_room`.

```{r}
# Your code here!

```

## time events

Using `extrasteps::step_time_event()` and the [almanac](https://davisvaughan.github.io/almanac/index.html) package, we can create useful time features.

```{r}
library(almanac)

rule_1 <- weekly() |>
  recur_on_weekdays() |>
  rsetdiff(hol_christmas())

rule_2 <- monthly(since = "2000-01-01") |>
  recur_on_interval(3) |>
  recur_on_day_of_month(1)

rule_3 <- yearly("1997-06-05") |>
  recur_on_day_of_week("Thursday") |>
  recur_on_month_of_year(c("Jun", "July", "Aug"))
```

## `step_time_event()`

Create a list of rules (last slide) and pass them to the `rules` argument of [`extrasteps::step_time_event()`](https://emilhvitfeldt.github.io/extrasteps/reference/step_time_event.html)

```{r}
#| eval: false
rules <- list(rule_1 = rule_1, rule_2 = rule_2, rule_3 = rule_3)

recipe(~arrival_date, data = hotel_rates) |>
  step_time_event(arrival_date, rules = rules)
```
