---
title: "Extras - Feature selection - Classwork"
subtitle: "Practical Machine Learning with tidymodels"
editor_options: 
  chunk_output_type: console
---

We recommend restarting R between each slide deck!

## Setup

```{r}
library(tidymodels)
library(important)
library(probably)
library(mirai)

tidymodels_prefer()
theme_set(theme_bw())
options(pillar.advice = FALSE, pillar.min_title_chars = Inf)
daemons(parallel::detectCores())
```

## More setup

```{r}
set.seed(429)
class_data <- 
  # Nonlinear classification with a low event rate
  sim_classification(5000, intercept = -10) |> 
  # Add some noise predictors that are correlated with one another
  bind_cols(sim_noise(5000, num_var = 15, cov_type = "toeplitz", cov_param = 0.5))

sim_split <- initial_split(class_data, prop = 0.75, strata = class)
sim_train <- training(sim_split)
sim_test  <- testing(sim_split)

set.seed(523)
sim_rs <- vfold_cv(sim_train, v = 10, strata = class)
```

## K-nearest neighbors

```{r}
rec <-
  recipe(class ~ ., data = sim_train) |>
  step_predictor_best(
    all_predictors(),
    score = "imp_rf",
    prop_terms = tune(),
    id = "filter"
  ) |>
  step_normalize(all_numeric_predictors())
  
knn_spec <- 
  nearest_neighbor(neighbors = tune(), weight_func = tune()) |> 
  set_mode("classification")
  
thrsh_tlr <-
  tailor() |>
  adjust_probability_threshold(threshold = tune()) 
```

## Setup the workflow

```{r}
knn_wflow <- workflow(rec, knn_spec, thrsh_tlr)

knn_param <-
  knn_wflow |>
  extract_parameter_set_dials() |>
  update(
    threshold = threshold(c(0.001, 0.1)),
    neighbors = neighbors(c(1, 50)),
    prop_terms = prop_terms(c(0.1, 1))
  )
```

## Tuning results

```{r}
cls_mtr <- metric_set(brier_class, roc_auc, sensitivity, specificity)
ctrl <- control_grid(save_pred = TRUE, save_workflow = TRUE)

set.seed(12)
knn_res <-
  knn_wflow |>
  tune_grid(
    resamples = sim_rs,
    grid = 20,
    control = ctrl,
    metrics = cls_mtr,
    param_info = knn_param
  )
```

## Grid results

```{r}
autoplot(knn_res)
```

## Brier results

```{r}
autoplot(knn_res, metric = "brier_class") + 
  facet_grid(. ~ name, scale = "free_x") 
```

## ROC curve results

```{r}
autoplot(knn_res, metric = "roc_auc") + 
  facet_grid(. ~ name, scale = "free_x") 
```

## Sensitivity/specificity results

```{r}
autoplot(knn_res, metric = c("sensitivity", "specificity"))
```

## Fit the model and get filter information

```{r}
knn_fit <- fit_best(knn_res, metric = "brier_class")
filter_info <-
	knn_fit |>
	extract_recipe() |>
	tidy(id = "filter")

filter_info
```

## Other steps

The important package has two other feature selection steps that can be used with multiple scores:

`step_predictor_retain()`: choose predictors based on a logical statement. Example:

```r
imp_rf > 2 & cor_pearson >= 0.75
```

`step_predictor_desirability()`: choose multiple scores to compute then use desirability functions to rank them:

```r
desirability(
  maximize(correlation),
  maximize(imp_rf)
)
```

## Manual approach

We already have our fitted model and, if we are happy with it:

```{r}
test_pred <- augment(knn_fit, sim_test)
test_pred |> cls_mtr(class, estimate = .pred_class, .pred_class_1)
```

## Checking (Approximate) Calibration

```{r}
test_pred |>
  cal_plot_windowed(
    truth = class,
    estimate = .pred_class_1,
    window_size = 0.2,
    step_size = 0.025,
  )
```

## Automated approach

We have to start with a finalized workflow (i.e., no `tune()` values):

```{r}
knn_best <- select_best(knn_res, metric = "brier_class")
knn_last_wflow <- finalize_workflow(knn_wflow, knn_best)
```

`last_fit()` uses the original split object to fit, predict, and measure the model using the test set:

```{r}
knn_test_res <- 
  knn_last_wflow |> 
  last_fit(sim_split, metrics = cls_mtr)
  
knn_test_res
```

We can pick out the parts that we want:

```{r}
knn_final_fit <- knn_test_res |> extract_workflow()
knn_test_pred <- knn_test_res |> collect_predictions()
knn_test_mtr  <- knn_test_res |> collect_metrics()

knn_test_mtr
```
