---
title: "Extras - Racing - Classwork"
subtitle: "Practical Machine Learning with tidymodels"
editor_options: 
  chunk_output_type: console
---

We recommend restarting R between each slide deck!

## Setup

```{r}
library(tidymodels)
library(finetune)
library(bonsai)

tidymodels_prefer()
theme_set(theme_bw())
options(pillar.advice = FALSE, pillar.min_title_chars = Inf)

cores <- parallel::detectCores() - 1
```

## More setup

```{r}
# Make our example data for this section
set.seed(429)
class_data <- 
  # Nonlinear classification data
  sim_classification(5000) |> 
  # Add some noise predictors that are correlated with one another
  bind_cols(sim_noise(5000, num_var = 15, cov_type = "toeplitz", cov_param = 0.5))

sim_split <- initial_split(class_data, prop = 0.75, strata = class)
sim_train <- training(sim_split)
sim_test  <- testing(sim_split)

set.seed(523)
sim_rs <- vfold_cv(sim_train, v = 10, strata = class)
```

## Boosted Tree Tuning Parameters

We'll need to load the bonsai package. This has the information needed to use lightgbm

```{r}
library(bonsai)

lgbm_spec <-
  boost_tree(
    trees = tune(),
    learn_rate = tune(),
    mtry = tune(),
    min_n = tune(),
    stop_iter = tune()
  ) |>
  set_mode("classification") |>
  # Turn off within-tree parallel processing; it's faster to run 
  # the resamples/configurations in parallel
  set_engine("lightgbm", num_threads = 1) 

# No preprocessing required:
lgbm_wflow <- workflow(class ~ ., lgbm_spec)
```

## Racing

```{r}
library(finetune)

# Set this to true to demo
ctrl <- control_race(verbose_elim = FALSE)

# Optimizes on the first metric in the set
cls_mtr <- metric_set(brier_class, roc_auc, sensitivity, specificity)

mirai::daemons(cores)

set.seed(321)
lgbm_res <-
  lgbm_wflow |>
  tune_race_anova(              # <- very similar syntax to tune_grid()
    resamples = sim_rs,
    # Let's use a larger grid
    grid = 50,
    control = ctrl,
    metrics = cls_mtr
  )
```

## Racing Results

```{r}
show_best(lgbm_res, metric = "brier_class")
```

`select_best()` never considers candidate models that did not get to the end of the race.

There is a helper function to see how candidate models were removed from consideration.

```{r}
plot_race(lgbm_res)
```

## Your turn

* Run `tune_race_anova()` with a different seed and/or a different metric.
* Did you get the same or similar results?

```{r}
# Your code here!

```
