---
title: "4 - Evaluating models - Classwork"
subtitle: "Practical Machine Learning with tidymodels"
editor_options: 
  chunk_output_type: console
---

We recommend restarting R between each slide deck!

## Setup

Setup from deck 3:

```{r}
library(tidymodels)
library(probably)

set.seed(123)
cls_split <- initial_split(cls_data_2026, prop = 0.8)
cls_train <- training(cls_split)
cls_test <- testing(cls_split)

tree_spec <- decision_tree(cost_complexity = 0.0001, mode = "classification")
cls_wflow <- workflow(class ~ ., tree_spec)
cls_fit <- fit(cls_wflow, cls_train)
```

## A logistic regression fit

```{r}
lr_fit <- 
  logistic_reg() |> 
  fit(class ~ ., data = cls_train)

augment(lr_fit, new_data = cls_train) |> print(n = 5)

lr_re_pred <- augment(lr_fit, new_data = cls_train)
```

## Confusion matrix

```{r}
lr_re_pred |>
  conf_mat(truth = class, estimate = .pred_class)
```

```{r}
lr_re_pred |>
  conf_mat(truth = class, estimate = .pred_class) |>
  autoplot(type = "heatmap")
```

## Metrics for model performance

```{r}
lr_re_pred |>
  accuracy(truth = class, estimate = .pred_class)
```

## Dangers of accuracy

We need to be careful of using `accuracy()` since it can give "good" performance by only predicting one way with imbalanced data:

```{r}
lr_re_pred |>
  mutate(.pred_class = factor("class_1", levels = c("class_1", "class_2"))) |>
  accuracy(truth = class, estimate = .pred_class)
```

## Brier score

```{r}
lr_re_pred |> 
  brier_class(truth = class, .pred_class_1) 
```

Smaller values are better.

## Your turn

The probably package was used to make the calibration plots:

```{r}
library(probably)
lr_re_pred |> cal_plot_breaks(class, .pred_class_1)
```

Try the other calibration curve methods in probably:

* `cal_plot_logistic()`
* `cal_plot_windowed()`

Do you prefer one over the others?

```{r}
# Your code here!

```

## Metrics for model performance

We can use `metric_set()` to combine multiple calculations into one

```{r}
cls_metrics <- metric_set(accuracy, brier_class)

lr_re_pred |>
  cls_metrics(truth = class, estimate = .pred_class, .pred_class_1)
```

Metrics and metric sets work with grouped data frames!

```{r}
lr_re_pred |>
  mutate(group = rep_len(LETTERS[1:3], nrow(cls_train))) |> 
  group_by(group) |>
  cls_metrics(truth = class, estimate = .pred_class, .pred_class_1)
```

## Dangers of overfitting

```{r}
lr_fit |>
  augment(cls_train) |> 
  print(n = 5)
```

We call this "resubstitution" or "repredicting the training set"

```{r}
lr_fit |>
  augment(cls_train) |>
  accuracy(class, .pred_class)
```

We call this a "resubstitution estimate"

Compare the training set with the test set:

```{r}
lr_fit |>
  augment(cls_test) |>
  accuracy(class, .pred_class)
```

Why aren't the numbers very different? Can basic logistic regression overfit?

⚠️ Remember that we're only repredicting the training set to demonstrate overfitting - it is not a valid way to evaluate a model.

⚠️ Don't use the test set until the *end* of your modeling analysis.

Same comparison, using the Brier score:

```{r}
lr_fit |>
  augment(cls_train) |>
  brier_class(class, .pred_class_1)
```

```{r}
lr_fit |>
  augment(cls_test) |>
  brier_class(class, .pred_class_1)
```

## Your turn

If we use 10 folds, what percent of the training data

* ends up in analysis
* ends up in assessment

## Cross-validation

```{r}
vfold_cv(cls_train) # v = 10 is default
```

What is in this?

```{r}
cls_folds <- vfold_cv(cls_train)
cls_folds$splits[1:3]
```

```{r}
vfold_cv(cls_train, v = 5)
```

We'll use this setup:

```{r}
set.seed(343)
cls_folds <- vfold_cv(cls_train, v = 10)
cls_folds
```

## Fit our model to the resamples

```{r}
lr_wflow <- workflow(class ~ ., logistic_reg())

cls_res <- fit_resamples(lr_wflow, cls_folds)
cls_res
```

## Evaluating model performance

```{r}
cls_res |>
  collect_metrics()
```

```{r}
cls_res |>
  collect_metrics() |> 
  select(.metric, mean, std_err, n)
```

## Your turn

Rerun the same code three times with different random number seeds.

How much do the results change?

```{r}
# Your code here!

```

## Evaluating model performance

```{r}
# Save the assessment set results
ctrl_rs <- control_resamples(save_pred = TRUE)
cls_res <- fit_resamples(lr_wflow, cls_folds, control = ctrl_rs)

cls_res
```

```{r}
# Save the assessment set results
cls_preds <- collect_predictions(cls_res)
cls_preds
```

```{r}
cls_preds |> 
  group_by(id) |>
  cls_metrics(truth = class, estimate = .pred_class, .pred_class_1)
```

## Where are the fitted models?

```{r}
cls_res
```

🗑️

## Create a random forest model

```{r}
rf_spec <- rand_forest(trees = 1000, mode = "classification")
rf_spec
```

```{r}
rf_wflow <- workflow(class ~ ., rf_spec)
rf_wflow
```

## Your turn

Use `fit_resamples()` and `rf_wflow` to:

* keep predictions
* compute metrics

```{r}
# Your code here!

```

## Evaluating model performance

```{r}
ctrl_rs <- control_resamples(save_pred = TRUE)

# Random forest uses random numbers so set the seed first

set.seed(2)
rf_res <- fit_resamples(rf_wflow, cls_folds, control = ctrl_rs)
collect_metrics(rf_res)
```

probably has a nice interface for calibration curves:

```{r}
cal_plot_windowed(rf_res, step_size = 0.03)
```

## Your turn

Repredict the training set using the random forest model.

Does the random forest model give overly optimistic results on the training set?

```{r}
# Your code here!

```

# Backup slides

## Metrics for model performance

```{r}
lr_re_pred |>
  sensitivity(truth = class, estimate = .pred_class)
```

```{r}
lr_re_pred |>
  specificity(truth = class, estimate = .pred_class)
```

## ROC curves

```{r}
# Assumes _first_ factor level is event; there are options to change that
lr_re_pred |> 
  roc_curve(truth = class, .pred_class_1) |>
  slice(1, 20, 50)

lr_re_pred |> 
  roc_auc(truth = class, .pred_class_1)
```

## ROC curve plot

```{r}
augment(lr_fit, 
        new_data = cls_train) |> 
  roc_curve(truth = class, 
            .pred_class_1) |>
  autoplot()
```

## Your turn

Compute and plot an ROC curve for your current model.

What data are being used for this ROC curve plot?

```{r}
# Your code here!

```

## Bootstrapping

```{r}
set.seed(3214)
bootstraps(cls_train)
```

## Your turn

Create:

* Monte Carlo Cross-Validation sets
* validation set

(use the reference guide to find the functions)

Don't forget to set a seed when you resample!

```{r}
# Your code here!

```

## Monte Carlo Cross-Validation

```{r}
set.seed(322)
mc_cv(cls_train, times = 10)
```

## Validation set

```{r}
set.seed(853)
cls_val_split <- initial_validation_split(cls_data_2026)
validation_set(cls_val_split)
```
