---
title: "Extras - Iterative search - Classwork"
subtitle: "Practical Machine Learning with tidymodels"
editor_options: 
  chunk_output_type: console
---

We recommend restarting R between each slide deck!

## Setup

```{r}
library(tidymodels)
library(important)
library(probably)

tidymodels_prefer()
theme_set(theme_bw())
options(pillar.advice = FALSE, pillar.min_title_chars = Inf)
mirai::daemons(parallel::detectCores())
```

## More setup

```{r}
# Make our example data for this section
set.seed(429)
class_data <- 
  # Nonlinear classification data
  sim_classification(5000) |> 
  # Add some noise predictors that are correlated with one another
  bind_cols(sim_noise(5000, num_var = 15, cov_type = "toeplitz", cov_param = 0.5))

set.seed(429)
sim_split <- initial_split(class_data, prop = 0.75, strata = class)
sim_train <- training(sim_split)
sim_test  <- testing(sim_split)

set.seed(523)
sim_rs <- vfold_cv(sim_train, v = 10, strata = class)
```

## Neural network

```{r}
nnet_spec <-
  mlp(hidden_units = tune(), penalty = tune(), learn_rate = tune(),
      epochs = 100, activation = tune()
  ) |>
  set_engine("brulee", stop_iter = 10) |>
  set_mode("classification")

rec <- 
  recipe(class ~ ., data = sim_train) |> 
  step_normalize(all_numeric_predictors())
  
thrsh_tlr <-
  tailor() |>
  adjust_probability_threshold(threshold = tune()) 
  
nnet_wflow <- workflow(rec, nnet_spec, thrsh_tlr)

nnet_param <-
  nnet_wflow |> 
  extract_parameter_set_dials() |> 
  update(threshold = threshold(c(0.0001, 0.1)))
  
cls_mtr <- metric_set(brier_class, roc_auc, sensitivity, specificity)
```

## Iterative Search

Suppose that we are only tuning the learning rate in our neural network. We could do something like:

```r
brier_model <- lm(brier ~ learn_rate, data = resample_results)
```

and use this to predict and rank new learning rate candidates.

## Your turn

Your GP makes predictions on two new candidate tuning parameters.

We want to minimize error.

Which should we choose?

## An Initial Grid

```{r}
set.seed(12)
init_res <-
  nnet_wflow |>
  tune_grid(
    resamples = sim_rs,
    grid = nrow(nnet_param) + 6, # for activation values + 1 extra
    param_info = nnet_param,
    metrics = cls_mtr
  )

show_best(init_res, metric = "brier_class", n = 3) |> select(-.metric, -.estimator)
```

## BO using tidymodels

```{r}
ctrl_bo <- control_bayes(verbose_iter = TRUE, no_improve = Inf) 

set.seed(125)
nnet_bayes_res <-
  nnet_wflow |>
  tune_bayes(
    resamples = sim_rs,
    initial = init_res,     # <- initial results
    iter = 25,
    control = ctrl_bo,
    param_info = nnet_param,
    metrics = cls_mtr
  )
```

## Best results

```{r}
show_best(nnet_bayes_res, metric = "brier_class") |> select(-.metric, -.estimator)
```

## Plotting BO Results

```{r}
autoplot(nnet_bayes_res, metric = "brier_class")
```

```{r}
autoplot(nnet_bayes_res, metric = "brier_class", type = "parameters")
```

```{r}
autoplot(nnet_bayes_res, metric = "brier_class", type = "performance")
```

## Your turn

Let's try a different acquisition function: `conf_bound(kappa)`.

We'll use the `objective` argument to set it.

Choose your own `kappa` value:

* Larger values will explore the space more.
* "Large" values are usually less than one.

**Bonus points**: Before the optimization is done, press `<esc>` and see what happens.

```{r}
# Your code here!

```
