---
title: "Extras - Postprocessing - Classwork"
subtitle: "Practical Machine Learning with tidymodels"
editor_options: 
  chunk_output_type: console
---

We recommend restarting R between each slide deck!

## Setup

```{r}
library(tidymodels)
library(desirability2)
library(probably)
library(mirai)

# check torch:
if (torch::torch_is_installed()) {
  library(torch)
}

tidymodels_prefer()
theme_set(theme_bw())
options(pillar.advice = FALSE, pillar.min_title_chars = Inf)
daemons(parallel::detectCores())
```

## More setup

```{r}
set.seed(429)
class_data <- 
  # Nonlinear classification with a low event rate
  sim_classification(5000, intercept = -10) |> 
  # Add some noise predictors that are correlated with one another
  bind_cols(sim_noise(5000, num_var = 15, cov_type = "toeplitz", cov_param = 0.5))

sim_split <- initial_split(class_data, prop = 0.75, strata = class)
sim_train <- training(sim_split)
sim_test  <- testing(sim_split)

set.seed(523)
sim_rs <- vfold_cv(sim_train, v = 10, strata = class)
```

## Our neural network model

```{r}
rec <- 
  recipe(class ~ ., data = sim_train) |> 
  step_normalize(all_numeric_predictors())

nnet_spec <- 
  mlp(hidden_units = tune(), penalty = tune(), learn_rate = tune(), 
      epochs = 100, activation = tune()) |> 
  # Remove the class_weights argument
  set_engine("brulee", stop_iter = 10) |> 
  set_mode("classification")
  
nnet_wflow <- workflow(rec, nnet_spec)

nnet_param <- 
  nnet_wflow |> 
  extract_parameter_set_dials()   
```

## Tailors

```{r}
thrsh_tlr <-
  tailor() |>
  adjust_probability_threshold(threshold = 1 / 3)

thrsh_tlr
```

## Manual use of a tailor

```{r}
three_rows <- 
  tribble(
     ~ class, ~ .pred_class, ~.pred_class_1, ~.pred_class_2,
   "class_1",     "class_1",          0.6,             0.4,
   "class_1",     "class_2",          0.4,             0.6, 
   "class_2",     "class_2",          0.1,             0.9  
  ) |> 
  mutate(across(where(is.character), factor))

thrsh_fit <-
  thrsh_tlr |>
  fit(
    three_rows,
    outcome = class,
    estimate = .pred_class,
    .pred_class_1:.pred_class_2  # No argument name and order matches factor levels
  )
```

`predict()` applies the adjustments:

```{r}
thrsh_fit

predict(thrsh_fit, three_rows)
```

## tailors within workflows

```{r}
nnet_wflow <- workflow(rec, nnet_spec, thrsh_tlr)
```

## Your turn

Discuss with those around you what the "ordering rules" could be.

## Tuning the probability threshold

```{r}
thrsh_tlr <-
  tailor() |>
  adjust_probability_threshold(threshold = tune())

nnet_thrsh_wflow <- workflow(rec, nnet_spec, thrsh_tlr)
  
nnet_thrsh_param <- 
  nnet_thrsh_wflow |> 
  extract_parameter_set_dials() |> 
  update(threshold = threshold(c(0.001, 0.5)))
```

Nearly the same code as before:

```{r}
ctrl <- control_grid(save_pred = TRUE, save_workflow = TRUE)
cls_mtr <- metric_set(brier_class, roc_auc, sensitivity, specificity)

set.seed(12)
nnet_thrsh_res <-
  nnet_thrsh_wflow |>
  tune_grid(
    resamples = sim_rs,
    grid = 25,
    param_info = nnet_thrsh_param, 
    control = ctrl,
    metrics = cls_mtr
  )
```

## Grid results

```{r}
autoplot(nnet_thrsh_res)
```

## Multimetric optimization

```{r}
nnet_thrsh_res |>
  show_best_desirability(
    maximize(sensitivity),
    minimize(brier_class),
    constrain(specificity, low = 0.8, high = 1.0)
  ) |>
  relocate(threshold, sensitivity, specificity, brier_class, .d_overall)
```

## Calibration

```{r}
more_sens <-
  nnet_thrsh_res |>
  select_best_desirability(
    maximize(sensitivity),
    minimize(brier_class),
    constrain(specificity, low = 0.8, high = 1.0)
  )

nnet_thrsh_res |>
  collect_predictions(
    parameters = more_sens
  ) |>
  cal_plot_windowed(
    truth = class,
    estimate = .pred_class_1,
    window_size = 0.2,
    step_size = 0.025,
  )
```

## Your turn

* Based on previous results, choose and fix a specific activation type (i.e., no `tune()`).
* Add a calibrator to your tailor with a `method = tune()` value.
* Run another grid search

Does it help with this data set?

```{r}
# Your code here!

```
