---
title: "Extras - Case Study on Transportation - Classwork"
subtitle: "Practical Machine Learning with tidymodels"
editor_options: 
  chunk_output_type: console
---

We recommend restarting R between each slide deck!

## Your turn

Take a look at these data for a few minutes and see if you can find any interesting characteristics in the predictors or the outcome.

```{r}
library(tidymodels)
library(rules)
data("Chicago")
dim(Chicago)
stations
```

```{r}
# Your code here!

```

## Splitting with Chicago data

Let's put the last two weeks of data into the test set. `initial_time_split()` can be used for this purpose:

```{r}
data(Chicago)

chi_split <- initial_time_split(Chicago, prop = 1 - (14/nrow(Chicago)))
chi_split

chi_train <- training(chi_split)
chi_test  <- testing(chi_split)

## training
nrow(chi_train)
 
## testing
nrow(chi_test)
```

## Times series resampling

Use the `date` column to find the date data.

```{r}
#| eval: false
chi_rs <-
  chi_train |>
  sliding_period(
    index = "date",  
  )
```

Our units will be weeks.

```{r}
#| eval: false
chi_rs <-
  chi_train |>
  sliding_period(
    index = "date",  
    period = "week",
  )
```

Every analysis set has 15 years of data.

```{r}
#| eval: false
chi_rs <-
  chi_train |>
  sliding_period(
    index = "date",  
    period = "week",
    lookback = 52 * 15  
  )
```

Every assessment set has 2 weeks of data.

```{r}
#| eval: false
chi_rs <-
  chi_train |>
  sliding_period(
    index = "date",  
    period = "week",
    lookback = 52 * 15,
    assess_stop = 2,
  )
```

Increment by 2 weeks so that there are no overlapping assessment sets.

```{r}
chi_rs <-
  chi_train |>
  sliding_period(
    index = "date",  
    period = "week",
    lookback = 52 * 15,
    assess_stop = 2,
    step = 2 
  )
```

```{r}
chi_rs$splits[[1]] |> assessment() |> pluck("date") |> range()
chi_rs$splits[[2]] |> assessment() |> pluck("date") |> range()
```

## Our resampling object

```{r}
chi_rs
```

## Feature engineering with recipes

```{r}
chi_rec <- 
  recipe(ridership ~ ., data = chi_train)
```

Based on the formula, the function assigns columns to roles of "outcome" or "predictor"

```{r}
summary(chi_rec)
```

## A recipe - work with dates

```{r}
chi_rec <- 
  recipe(ridership ~ ., data = chi_train) |> 
  step_date(date, features = c("dow", "month", "year")) 
```

Add indicators for major holidays.

```{r}
chi_rec <- 
  recipe(ridership ~ ., data = chi_train) |> 
  step_date(date, features = c("dow", "month", "year")) |> 
  step_holiday(date) 
```

```{r}
chi_rec <- 
  recipe(ridership ~ ., data = chi_train) |> 
  step_date(date, features = c("dow", "month", "year")) |> 
  step_holiday(date) |> 
  update_role(date, new_role = "id") |>
  update_role_requirements(role = "id", bake = TRUE)
```

## A recipe - remove constant columns

```{r}
chi_rec <- 
  recipe(ridership ~ ., data = chi_train) |> 
  step_date(date, features = c("dow", "month", "year")) |> 
  step_holiday(date) |> 
  update_role(date, new_role = "id") |>
  update_role_requirements(role = "id", bake = TRUE) |> 
  step_zv(all_nominal_predictors()) 
```

## A recipe - handle correlations

```{r}
chi_pca_rec <- 
  chi_rec |> 
  step_normalize(all_of(!!stations)) |> 
  step_pca(all_of(!!stations), num_comp = tune())
```

## Make some models

```{r}
cb_spec <- cubist_rules(committees = 25, neighbors = tune())
mars_spec <- mars(prod_degree = tune()) |> set_mode("regression")
lm_spec <- linear_reg()

chi_set <- 
  workflow_set(
    list(pca = chi_pca_rec, basic = chi_rec), 
    list(cubist = cb_spec, mars = mars_spec, lm = lm_spec)
  ) |> 
  # Evaluate models using mean absolute errors
  option_add(metrics = metric_set(mae))
```

## Process them on the resamples

```{r}
# Set up some objects for stacking ensembles (in a few slides)
grid_ctrl <- control_grid(save_pred = TRUE, save_workflow = TRUE)

chi_res <- 
  chi_set |> 
  workflow_map(
    resamples = chi_rs,
    grid = 10,
    control = grid_ctrl,
    verbose = TRUE,
    seed = 12
  )
```

## How do the results look?

```{r}
rank_results(chi_res)
```

## Plot the results

```{r}
autoplot(chi_res)
```

## Pull out specific results

```{r}
chi_res |> 
  extract_workflow_set_result("pca_cubist") |> 
  autoplot()
```

## Building a model stack

```{r}
library(stacks)
```

## Start the stack and add members

```{r}
chi_stack <- 
  stacks() |> 
  add_candidates(chi_res)
```

## Estimate weights for each candidate

```{r}
set.seed(122)
chi_stack_res <- blend_predictions(chi_stack, penalty = 10^seq(-6, -1, length.out = 25))

chi_stack_res
```

## How did it do?

```{r}
autoplot(chi_stack_res)
```

## What does it use?

```{r}
autoplot(chi_stack_res, type = "weights")
```

## Fit the required candidate models

```{r}
chi_stack_res <- fit_members(chi_stack_res)
```

## The test set: best Cubist model

```{r}
best_cubist <- 
  chi_res |> 
  extract_workflow_set_result("pca_cubist") |> 
  select_best()

cubist_res <- 
  chi_res |> 
  extract_workflow("pca_cubist") |> 
  finalize_workflow(best_cubist) |> 
  last_fit(split = chi_split, metrics = metric_set(mae))
```

## The test set: stack ensemble

```{r}
stack_pred <- 
  predict(chi_stack_res, chi_test) |> 
  bind_cols(chi_test)
```

## Compare the results

```{r}
collect_metrics(cubist_res)

stack_pred |> mae(ridership, .pred)
```

## Plot the test set

```{r}
library(probably)
cubist_res |> 
  collect_predictions() |> 
  ggplot(aes(ridership, .pred)) + 
  geom_point(alpha = 1 / 2) + 
  geom_abline(lty = 2, col = "green") + 
  coord_obs_pred()
```
