---
title: "2 - Your data budget - Classwork"
subtitle: "Introduction to tidymodels"
editor_options: 
  chunk_output_type: console
---

We recommend restarting R between each slide deck!

## Data on forests in Washington

```{r}
library(tidymodels)
library(forested)

forested
```

## Your turn

When is a good time to split your data?

## Data splitting and spending

```{r}
set.seed(123)

forested_split <- initial_split(forested)
forested_split
```

Extract the training and testing sets

```{r}
forested_train <- training(forested_split)
forested_test <- testing(forested_split)
```

## Validation set

```{r}
set.seed(123)
initial_validation_split(forested, prop = c(0.6, 0.2))
```

## Your turn

Split your data so 20% is held out for the test set.

Try out different values in `set.seed()` to see how the results change.

Hint: Which argument in `initial_split()` handles the proportion split into training vs testing?

```{r}
# Your code here!

```

## Your turn

Explore the `forested_train` data on your own!

* What's the distribution of the outcome, `forested`?
* What's the distribution of numeric variables like `precip_annual`?
* How does the distribution of `forested` differ across the categorical variables?

```{r}
# Your code here!

```
