---
title: "3 - What makes a model? - Classwork"
subtitle: "Introduction to tidymodels"
editor_options: 
  chunk_output_type: console
---

We recommend restarting R between each slide deck!

## Setup

Setup from deck 2

```{r}
library(tidymodels)
library(forested)

set.seed(123)
forested_split <- initial_split(forested, prop = 0.8)
forested_train <- training(forested_split)
forested_test <- testing(forested_split)
```

## Your turn

How do you fit a linear model in R?

How many different ways can you think of?

Discuss with your neighbor!

## To specify a model

```{r}
# Model
linear_reg()

# Engine
linear_reg() %>%
  set_engine("glmnet")

# Mode - Some models have a default mode, others don't
decision_tree() %>% 
  set_mode("regression")
```

## Your turn

Edit the chunk below to use a logistic regression model.

Extension/Challenge: Edit this code to use a different model. For example, try using a conditional inference tree as implemented in the partykit package by changing the engine - or try an entirely different model type!

All available models are listed at <https://www.tidymodels.org/find/parsnip/>

```{r tree_spec}
tree_spec <- decision_tree() %>% 
  set_mode("classification")

tree_spec
```

## A model workflow

```{r}
tree_spec <-
  decision_tree() %>% 
  set_mode("classification")
```

Fit parsnip specification:

```{r}
tree_spec %>% 
  fit(forested ~ ., data = forested_train) 
```

Fit with a workflow:

```{r}
workflow() %>%
  add_formula(forested ~ .) %>%
  add_model(tree_spec) %>%
  fit(data = forested_train) 
```

"Shortcut" by specifying the preprocessor and model spec directly in the `workflow()` call:

```{r}
workflow(forested ~ ., tree_spec) %>% 
  fit(data = forested_train) 
```

## Your turn

Edit the chunk below to make a workflow with your own model of choice!

Extension/Challenge: Other than formulas, what kinds of preprocessors are supported?

```{r tree_wflow}
tree_spec <-
  decision_tree() %>% 
  set_mode("classification")

tree_wflow <- workflow() %>%
  add_formula(forested ~ .) %>%
  add_model(tree_spec)

tree_wflow
```

## Predict with your model

```{r}
tree_fit <-
  workflow(forested ~ ., tree_spec) %>% 
  fit(data = forested_train) 
```

## Your turn

What do you get from running the following code? What do you notice about the structure of the result?

```{r}
predict(tree_fit, new_data = forested_test)
```

## Your turn

What do you get from running the following code? How is `augment()` different from `predict()`?

```{r}
augment(tree_fit, new_data = forested_test)
```

## Understand your model

```{r}
library(rpart.plot)

tree_fit %>%
  extract_fit_engine() %>%
  rpart.plot(roundint = FALSE)
```

## Your turn

Try extracting the model engine object from your fitted workflow!

```{r}
### Your code here

```

What kind of object is it? What can you do with it?

⚠️ Never `predict()` with any extracted components!

You can also read the documentation for object extraction:
https://workflows.tidymodels.org/reference/extract-workflow.html

