---
title: "2 - Your data budget - Classwork"
subtitle: "Introduction to tidymodels"
editor_options: 
  chunk_output_type: console
---

We recommend restarting R between each slide deck!

## Data on taxi trips in Chicago in 2022

```{r}
library(tidymodels)

taxi
```

## Your turn

When is a good time to split your data?

## Data splitting and spending

```{r}
set.seed(123)

taxi_split <- initial_split(taxi)
taxi_split
```

Extract the training and testing sets

```{r}
taxi_train <- training(taxi_split)
taxi_test <- testing(taxi_split)
```

## Validation set

```{r}
set.seed(123)
initial_validation_split(taxi, prop = c(0.6, 0.2))
```

## Your turn

Split your data so 20% is held out for the test set.

Try out different values in `set.seed()` to see how the results change.

Hint: Which argument in `initial_split()` handles the proportion split into training vs testing?

```{r}
# Your code here!

```

## Your turn

Explore the `taxi_train` data on your own!

- What's the distribution of the outcome, tip?
- What's the distribution of numeric variables like distance?
- How does tip differ across the categorical variables?

```{r}
# Your code here!

```

## Stratification

```{r}
set.seed(123)

taxi_split <- initial_split(taxi, prop = 0.8, strata = tip)
taxi_train <- training(taxi_split)
taxi_test <- testing(taxi_split)
```
