---
title: "5 - Tuning models - Classwork"
subtitle: "Practical Machine Learning with tidymodels"
editor_options:
  chunk_output_type: console
---

We recommend restarting R between each slide deck!

## Setup

Setup from deck 4:

```{r}
library(tidymodels)
library(probably)

set.seed(123)
cls_split <- initial_split(cls_data_2026, prop = 0.8)
cls_train <- training(cls_split)
cls_test <- testing(cls_split)

set.seed(343)
cls_folds <- vfold_cv(cls_train, v = 10)
```

## Your turn

Open the help page for `boost_tree()` (via `?boost_tree`).

Look at the *engine documentation* for xgboost models.

```{r}
# Your code here!

```

## Specifying the model

```{r}
# xgboost is the default engine
bst_spec <- boost_tree(
  trees = 500,
  learn_rate = tune(),
  mode = "classification"
)
bst_wflow <- workflow(class ~ ., bst_spec)

bst_wflow
```

## Working with Parameters

```{r}
mtry_param <-
  boost_tree(mtry = tune()) |>
  extract_parameter_set_dials()
mtry_param

mtry_param |>
  update(mtry = mtry(c(1, 10)))
```

## Try out multiple values

```{r}
set.seed(22)
bst_res <- tune_grid(
  bst_wflow,
  cls_folds,
  grid = 15
)
```

## Compare results

```{r}
autoplot(bst_res)
```

## Saving predictions

Let's repeat this process but make sure that we keep the out-of-sample
predictions

```{r}
set.seed(22)
bst_res <- tune_grid(
  bst_wflow,
  cls_folds,
  grid = 15,
  control = control_grid(save_pred = TRUE)
)
```

## Checking (Approximate) Calibration

```{r}
#| eval: false

bst_res |>
  cal_plot_windowed(
    window_size = 0.2,
    step_size = 0.025,
  )
```

## Compare results

Which is numerically best?

```{r}
show_best(bst_res, metric = "brier_class")

best_parameter <- select_best(bst_res, metric = "brier_class")
best_parameter
```

## Running in parallel

We can use the future or mirai packages to do this:

```{r}
cores <- parallelly::availableCores(logical = FALSE)
```

```{r}
#| eval: false

library(future)
plan(multisession, workers = cores)

# Now call `tune_grid()`!
```

```{r}
#| eval: false

library(mirai)
daemons(cores)

# Now call `tune_grid()`!
```

## Your turn

Modify your model workflow to tune additional parameters.

Use grid search to find the best parameter(s).

```{r}
# Your code here!

```

## The final fit

```{r}
bst_final_wflow <- finalize_workflow(bst_wflow, best_parameter)
# cls_split has train + test info
set.seed(690)
final_res <- last_fit(bst_final_wflow, cls_split)

final_res
```

## What is in `final_res`?

```{r}
collect_metrics(final_res)
```

These are metrics computed with the **test** set

```{r}
collect_predictions(final_res)
```

```{r}
final_res |>
  cal_plot_windowed(window_size = 0.2, step_size = 0.025) +
  geom_rug()
```

```{r}
extract_workflow(final_res)
```

Use this for **prediction** on new data, like for deploying
